### 2017-09-07 analysis/1-methods-toys.Rmd
added method plotting

---
title: "1-methods-toys"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Testing methods on simple toy datasets

```{r}
library(tidyverse)
library(dyneval)

set.seed(1)

tasks <- generate_toy_datasets()
tasks <- tasks[1, ]
```

We choose certain parameters for which we know that the method will find (or focus on finding) a linear trajectory, and will do this quickly
```{r}
method_params <- list(
  waterfall=list(),
  scorpius=list(),
  slingshot=list(),
  gpfates=list(nfates=1),
  stemid=list(clustnr=10, bootnr=10, pdishuf=10),
  tscan=list(),
  embeddr=list(nn_pct = 2),
  celltree_gibbs=list(sd_filter = 0),
  celltree_maptpx=list(sd_filter = 0),
  celltree_vem=list(sd_filter = 0),
  scuba=list(),
  slicer = list(min_branch_len=50),
  monocle_ddrtree=list(),
  wishbone = list(branch=F),
  pseudogp = list(iter=100, initialise_from="principal_curve"),
  mpath = list(numcluster=6),
  random_linear=list()
)
method_names <- names(method_params)

method_descriptions <- map(method_names, ~get(paste0("description_", .))()) %>% set_names(method_names)

metric_names <- c("mean_R_nx", "auc_R_nx", "Q_local", "Q_global", "correlation", "isomorphic", "robbie_network_score")
```

```{r}
results <- purrr::map(method_names, function(method_name) {
  library(dyneval)

  cat("Processing ", method_name, "\n", sep="")
  method <- method_descriptions[[method_name]]
  params <- method_params[[method_name]]

  factory <- function(fun) {
    function(...) {
      warn <- err <- NULL
      res <- withCallingHandlers(
        tryCatch(fun(...), error=function(e) {
          err <<- conditionMessage(e)
          NULL
        }), warning=function(w) {
          warn <<- append(warn, conditionMessage(w))
          invokeRestart("muffleWarning")
        })
      list(res, warn=warn, err=err, method_name=method_name)
    }
  }

  factory(function() {
    dyneval:::execute_evaluation(tasks, method, params, metrics = metric_names, suppress_output = F)
  }
  )()
})
```

Gather the output of the different methods, and check whether there were any errors
```{r}
scores <- purrr::map(results, function(.) {
  bind_cols(
    tibble(error=list(.$err), warning=list(.$warning), model=list(attr(.[[1]], "extras")$.models[[1]])),
    attr(.[[1]], "extras")$.summary,
    tibble(method_name=.$method_name)
  )
}) %>% bind_rows() %>% mutate(errored = !map_lgl(error, is.null), method_id=method_names) %>% dplyr::select(-contains("method_name1"))

scores %>% saveRDS("data/derived_data/toy_method_scores.rds")
```

```{r}
scores %>% filter(errored)

print(scores$error)
```

```{r}
scores %>%
  dplyr::select(-starts_with("time")) %>%
  gather(score_id, score, -method_name, -method_short_name, -task_id, -error, -warning) %>%
  ggplot() + geom_bar(aes(method_name, score, fill=score_id), stat="identity") + facet_wrap(~score_id) + coord_flip()
```