### 2017-09-12 analysis/method-qc.Rmd
added method qc

---
title: "method-qc"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(googlesheets)
library(cowplot)
rules_info <- gs_title("Trajectory Inference Methods") %>% gs_read(skip=1)
```

```{r}
ggplot(rules_info,aes(log(Citations+1),`QC score`))+
  geom_smooth(method='lm',formula=y~x) +
  geom_point() +
  ggrepel::geom_label_repel(aes(label=Name))

ggplot(rules_info,aes(log(Citations+1),`QC score`))+
  geom_smooth(method='lm',formula=y~x) +
  geom_point() +
  ggrepel::geom_label_repel(aes(label=Name))

rules_info$date <- pmax(
  lubridate::as_date(lubridate::ymd(rules_info$PubDate)),
  lubridate::as_date(lubridate::ymd(rules_info$Preprint)),
  na.rm=T
)

ggplot(rules_info,aes(date,`QC score`))+
  geom_smooth(method='lm',formula=y~x) +
  geom_point() +
  ggrepel::geom_label_repel(aes(label=Name))

rules_info$published <- !is.na(rules_info$PubDate)

ggplot(rules_info,aes(published,`QC score`))+
  geom_violin() +
  geom_point() +
  ggrepel::geom_text_repel(aes(label=Name))

```