### 2017-09-14 analysis/analyses/score_rules/2-rules.Rmd

---
title: "2-rules"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
scores <- read_rds("scores.rds")
toys <- read_rds("toys.rds")

# summarise the scores
# get gs-toy score
# get difference and fractions between gs-toy and gs-gs
scores_summary <- scores %>% gather(score_id, score, -toy_id, -comparison) %>%
  spread(comparison, score) %>% mutate(
    score=`gs-toy`,
    diff= `gs-toy`-`gs-gs`,
    frac=`gs-toy`/`gs-gs`
  ) %>% select(-`gs-toy`, -`gs-gs`) %>% left_join(toys, by="toy_id")

scores_summary %>%
  ggplot() +
  geom_boxplot(aes(toy_category, score, color=perturbator_id)) +
  facet_wrap(~score_id) +
  coord_flip()

scores_summary %>%
  ggplot() +
  geom_boxplot(aes(toy_category, diff, color=perturbator_id)) +
  facet_wrap(~score_id) +
  coord_flip()
```

```{r}
library(googlesheets)
rules_info <- gs_title("Metrics rules") %>% gs_read() %>% mutate(id=factor(id, levels=id))

source("rules.R")
rules_info$check_func <- map(paste0("check_", rules_info$id), get)

rules <- map2(
  rules_info$check_func,
  rules_info$id,
  ~.x(scores_summary) %>%
    mutate(rule_id=.y)
  ) %>% bind_rows() %>% left_join(rules_info, by=c("rule_id"="id"))

## Plot all rules
rules %>% ggplot(aes(rule_id, score_id)) +
  geom_raster(aes(fill = c("red", "green")[as.numeric(rule)+1])) +
  geom_text(aes(label = c("▼", "▲")[as.numeric(rule)+1]), color="white") +
  scale_fill_identity() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
## Possible score combinations satisfying all rules
allscores <-unique(rules$score_id)
score_ids_combinations <- map(1:length(allscores), function(x) map(combn(length(allscores), x, simplify = FALSE), ~allscores[.])) %>% unlist(recursive=FALSE)

combination_scores <- map(score_ids_combinations, function(score_ids_combination) {
  rules %>% filter(score_id %in% score_ids_combination) %>%
    group_by(rule_id) %>%
    summarise(any=any(rule), number=sum(rule, na.rm=TRUE)) %>%
    group_by() %>%
    summarise(all=all(any), avg_rules_retrieved=sum(number)/length(score_ids_combination), nscores = length(score_ids_combination)) %>%
    mutate(score_ids_combination=list(score_ids_combination))
}) %>% bind_rows()

combination_scores %>% filter(all) %>% arrange(nscores, -avg_rules_retrieved) %>% head(5)

combination_scores %>% filter(all) %>% arrange(nscores, -avg_rules_retrieved) %>% pull(score_ids_combination) %>% head(5)
```

## Aggregating metrics to get one final metric following all scores

In the end, we want to compare methods using one aggregated number which follows all the rules of a good evaluation metric. We will aggregate all combinations of metrics using different aggregation functions.
```{r}
score_ids_combinations <- map(1:length(allscores), function(x) map(combn(length(allscores), x, simplify = FALSE), ~allscores[.])) %>% unlist(recursive=FALSE)
new_scores_combinations <- tibble(
  combination = score_ids_combinations, 
  nscores = map_int(combination, length),
  combination_id = map_chr(combination, ~paste0(., collapse=","))
)
new_scores_aggregations <- tibble(
  aggregation_id = c("harmonic_mean", "arithmetic_mean", "geometric_mean", "max", "min"),
  aggregation_func = list(
    function(x) {if(any(x == 0)) {return(0)} else {length(x) /sum(1/x)}}, 
    mean, 
    function(x) {if(any(x == 0)) {return(0)} else {prod(x)^(1/length(x))}},
    max,
    min
  )
)

new_scores <- merge(new_scores_combinations, new_scores_aggregations) %>% as_tibble() %>% 
  mutate(score_id = paste0(aggregation_id, "(", combination_id, ")"))
```

```{r}
scores2 <- new_scores %>% rowwise() %>% do(
  tibble(score = apply(scores[.$combination], 1, .$aggregation_func), score_id=.$score_id, i=seq_len(nrow(scores)))
) %>% spread(score_id, score) %>% bind_cols(scores %>% select(toy_id, comparison))

scores2_summary <- scores2 %>% gather(score_id, score, -toy_id, -comparison) %>%
  spread(comparison, score) %>% mutate(
    score=`gs-toy`,
    diff= `gs-toy`-`gs-gs`,
    frac=`gs-toy`/`gs-gs`
  ) %>% select(-`gs-toy`, -`gs-gs`) %>% left_join(toys, by="toy_id")

rules <- map2(
  rules_info$check_func,
  rules_info$id,
  ~.x(scores2_summary) %>%
    mutate(rule_id=.y)
  ) %>% bind_rows() %>% left_join(rules_info, by=c("rule_id"="id"))

rules %>% group_by(score_id) %>% 
  summarise(allpassed = all(rule), npassed=sum(rule)) %>% 
  left_join(new_scores, by="score_id") %>% 
  filter(allpassed) %>% 
  arrange(nscores)
```