### 2017-09-17 analysis/method-qc.Rmd
add support for number of citations

---
title: "method-qc"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(googlesheets)
library(cowplot)
library(tidyverse)
rules_info <- gs_title("Trajectory Inference Methods") %>% 
  gs_read(ws = "Software", col_types = cols(GScholarClusterID = "c"), skip = 1)
```

Update number of citations
```{r}
# script_file <- "analysis/scholar.py"
# download.file("https://raw.githubusercontent.com/ckreibich/scholar.py/master/scholar.py", destfile = script_file)

num_citations_by_clusterid <- function(clusterid, scholar_file = script_file) {
  tryCatch({
    command <- paste0("python ", scholar_file, " -C ", clusterid, " --csv-header")
    output <- system(command, intern = T)
    tab <- readr::read_delim(paste(gsub("\n", " ", output), collapse = "\n"), delim = "|")
    sum(tab$num_citations)
  }, error = function(e) NA)
}

# If you're connected to UGentNet:
# rules_info <- rules_info %>%
#   mutate(Citations = pbapply::pbsapply(GScholarClusterID, num_citations_by_clusterid))
```

```{r}
ggplot(rules_info,aes(log(Citations+1),`QC score`))+
  geom_smooth(method='lm',formula=y~x) +
  geom_point() +
  ggrepel::geom_label_repel(aes(label=Name))

cit_breaks <- c(0, 1, 10, 100, max(rules_info$Citations))
ggplot(rules_info,aes(Citations+1,`QC score`))+
  geom_smooth(method='lm',formula=y~x) +
  geom_point() +
  ggrepel::geom_label_repel(aes(label=Name)) +
  scale_x_log10(breaks = cit_breaks+1, labels = cit_breaks)

rules_info$date <- pmax(
  lubridate::as_date(lubridate::ymd(rules_info$PubDate)),
  lubridate::as_date(lubridate::ymd(rules_info$Preprint)),
  na.rm=T
)

ggplot(rules_info,aes(date,`QC score`))+
  geom_smooth(method='lm',formula=y~x) +
  geom_point() +
  ggrepel::geom_label_repel(aes(label=Name)) +
  scale_y_continuous(breaks = seq_len(10))

rules_info$published <- !is.na(rules_info$PubDate)

ggplot(rules_info,aes(published,`QC score`))+
  geom_violin() +
  geom_point() +
  ggrepel::geom_text_repel(aes(label=Name))

```

```{r}

```