### 2018-01-16 analysis/analyses/3-score_rules/flip_score_characterisation.Rmd
added flip score characterisation

---
title: "Flip score characterisation"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dynmethods)
library(dyneval)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
models <- eval(formals(dyntoy:::generate_toy_milestone_network)$model)
models <- models[models != "BA"]
networks_synthetic <- map(models, dyntoy:::generate_toy_milestone_network) %>% tibble(milestone_network = ., model = models)
```

Add SCORPIUS predictions
```{r}
descriptions <- list(description_scorpius(), description_dpt())
data <- dyntoy::generate_toy_datasets(c("linear"), num_replicates=1, num_cells=30, num_genes=10) %>% dynutils::extract_row_to_list(1)

results <- map(descriptions, ~.$run_fun(data$counts))

networks_methods <- tibble(model=map_chr(descriptions, "name"), milestone_network=map(results, "milestone_network"))

```

```{r}
networks <- bind_rows(networks_synthetic, networks_methods)
design <- crossing(networks, networks)
```

```{r fig.height=7}
scores <- design %>% as.list() %>% pmap(function(milestone_network, milestone_network1, model, model1, ...) {
  # print(model)
  # print(model1)

  start <- Sys.time()
  score <- dyneval:::calculate_edge_flip(milestone_network, milestone_network1)
  time <- Sys.time() - start
  tibble(score=score, time=time)
}) %>% bind_rows()
design <- bind_cols(design, scores)

design %>% ggplot() + geom_raster(aes(model, model1, fill=score)) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
library(ggraph)
library(tidygraph)

networks$plot <- networks$milestone_network %>% map(function(net) {
  net %>% as_tbl_graph(directed=FALSE) %>%
    ggraph() +
      geom_edge_fan() + geom_node_point() + ggraph::theme_graph(plot_margin = margin(0, 0, 0, 0))
})

design %>%
  group_by(model) %>%
  arrange(-score) %>%
  left_join(networks %>% select(-milestone_network), by=c("model1"="model")) %>%
  summarise(plots = list(map2(plot, score, ~.x + ggtitle(round(.y, 2))))) %>%
  pull(plots) %>% map(~cowplot::plot_grid(plotlist=.))
```