### 2018-02-06 analysis/analyses/3-score_rules/flip_score_characterisation.Rmd

---
title: "Flip score characterisation"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dynmethods)
library(dyneval)
library(dynalysis)

experiment("score_rules/edge_flip")

knitr::opts_chunk$set(echo = TRUE)
```

Generate different trajectory topologies
```{r}
topologies <- eval(formals(dyntoy:::generate_toy_milestone_network)$model)
topologies <- topologies[topologies != "BA"]
networks_synthetic <- map(topologies, dyntoy:::generate_toy_milestone_network) %>% tibble(milestone_network = ., topology = topologies)
```

Add SCORPIUS predictions
```{r}
descriptions <- list(description_scorpius())
data <- dyntoy::generate_toy_datasets(c("linear"), num_replicates=1, num_cells=30, num_genes=10) %>% dynutils::extract_row_to_list(1)

results <- map(descriptions, ~.$run_fun(data$counts))

networks_methods <- tibble(topology=map_chr(descriptions, "name"), milestone_network=map(results, "milestone_network"))
```

```{r}
networks <- bind_rows(networks_synthetic, networks_methods)
design <- crossing(networks, networks)
```

```{r fig.height=7}
scores <- design %>% as.list() %>% pmap(function(milestone_network, milestone_network1, topology, topology1, ...) {
  # print(topology)
  # print(topology1)

  start <- Sys.time()
  score <- dyneval:::calculate_edge_flip(milestone_network, milestone_network1)
  time <- Sys.time() - start
  tibble(score=score, time=time)
}) %>% bind_rows()
design <- bind_cols(design, scores)

design %>% 
  ggplot() + 
  geom_raster(aes(topology, topology1, fill=score)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
exclude_topologies <- c("SCORPIUS", "linear_long")
edge_flip_comparison <- design %>% 
  filter(!topology %in% exclude_topologies, !topology1 %in% exclude_topologies) %>% 
  mutate(topology = factor(topology, levels=topologies)) %>% 
  mutate(topology1 = factor(topology1, levels=rev(topologies))) %>% 
  ggplot(aes(topology, topology1)) + 
  geom_tile(aes(fill=score)) + 
  geom_text(
    aes(
      label=round(score, 2) %>% gsub("0\\.", "\\.", .),
      color = ifelse(score > 0.5, "white", "black")
      )
  ) +
  viridis::scale_fill_viridis("Edge flip score", option="A", direction = -1) +
  scale_color_identity() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete("", labels=label_long, position="top", expand=c(0, 0)) +
  scale_y_discrete("", labels=label_long, expand=c(0, 0)) +
  theme(axis.text.x.top = element_text(hjust=0)) +
  coord_equal()
edge_flip_comparison
write_rds(edge_flip_comparison, figure_file("edge_flip_comparison.rds"))
```

```{r}
library(ggraph)
library(tidygraph)

networks$plot <- networks$milestone_network %>% map(function(net) {
  net %>% as_tbl_graph(directed=FALSE) %>%
    ggraph() +
      geom_edge_fan() + geom_node_point() + ggraph::theme_graph(plot_margin = margin(0, 0, 0, 0))
})

design %>%
  group_by(topology) %>%
  arrange(-score) %>%
  left_join(networks %>% select(-milestone_network), by=c("topology1"="topology")) %>%
  summarise(plots = list(map2(plot, score, ~.x + ggtitle(round(.y, 2))))) %>%
  pull(plots) %>% map(~cowplot::plot_grid(plotlist=.))
```