### 2018-03-05 analysis/paper/paper.Rmd
update rmd

---
title: "A comparison of single-cell trajectory inference methods: towards more accurate and robust tools"
bibliography: references.bib
csl: nature-biotechnology.csl
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage[table]{xcolor}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{tabu}
  - \usepackage[normalem]{ulem}
  - \usepackage{lscape}
  - \usepackage{pbox}
  - \usepackage[labelformat=empty]{caption}
  - \ExecuteBibliographyOptions{sorting=none,sortcites=true,autocite=superscript}
mainfont: Open Sans

---

*`r format(Sys.time(), '%d %B, %Y')`*
<!--
Tips for working in this document:
* Go to tools -> perferences and remove Use smart quotes
* Enable automatic zotero syncing to bibtex
* Make sure to pin the zotero bibtex references after first using them!

TITLES: 
Not all trajectory inference methods are created equal
Trajectory inference: work in progress
Trajectory inference: we’re not quite there yet
Paving the way for trajectories
Benchmarking trajectory inference methods
Bringing single-cell trajectory inference methods to the next level
Resolving disarray amongst trajectory inference methods with a comprehensive evaluation
A comparison of trajectory inference methods
Trajectory inference 2.0
We’re on the trajectory to better trajectory inference methods
On the road to more accurate and robust single-cell trajectory inference methods
Towards more accurate and robust single-cell trajectory inference methods
On the trajectory to more accurate,robust and trajectory-like single-cell trajectory inference method trajectories
A comparison of single-cell trajectory inference methods: towards more accurate and robust tools
-->

```{r setup, include=FALSE}
source(paste0(dynalysis::get_dynalysis_folder(), "/analysis/paper/setup.R"))
```
`r glue::collapse(sample(c("Wouter Saelens* ¹ ²", "Robrecht Cannoodt* ¹ ² ³")), ", ")`, Helena Todorov¹ ² ⁴ ⁵, Yvan Saeys¹ ²

\* Equal contribution  
¹ Data mining and Modelling for Biomedicine, VIB Center for Inflammation Research, Ghent, Belgium.  
² Department of Applied Mathematics, Computer Science and Statistics, Ghent University, Ghent, Belgium.  
³ Center for Medical Genetics, Ghent University Hospital, Ghent, Belgium.  
⁴ Centre International de Recherche en Infectiologie, Inserm, U1111, Université Claude Bernard Lyon 1, CNRS, UMR5308, École Normale Supérieure de Lyon, Univ Lyon, F-69007, LYON, France

# Abstract

Using single-cell transcriptomics or proteomics data, it is now possible to computationally order cells along trajectories, allowing the unbiased transcriptome-wide study of cellular dynamic processes. Since 2014, more than 50 trajectory inference methods have been developed, each with its own set of methodological characteristics. As a result, choosing a method to infer trajectories is often challenging, since a comprehensive assessment of the performance and robustness of each method is still lacking. In order to facilitate the comparison of the results of these methods to each other and to a gold standard, we developed a global framework to benchmark trajectory inference tools. . Using this framework, we compared the results from a total of  `r sum(methods$evaluated)` trajectory inference methods, on a large collection of real and synthetic datasets. We evaluate methods using several metrics, including accuracy of the inferred ordering, correctness of the network topology, code quality and user friendliness.

To our knowledge, this is the first comprehensive assessment of trajectory inference methods. For now, we exclusively evaluated the methods on their default parameters, but plan to add a detailed parameter tuning procedure in the future.  We gladly welcome any discussion and feedback on key decisions made as part of this study, including the metrics used in the benchmark, the quality control checklist, and the implementation of the method wrappers. These discussions can be held at [github.com/dynverse/dynverse/issues](https://www.github.com/dynverse/dynverse/issues).

# Introduction
<!-- §1: Single-cell omics and TI --->
Single-cell -omics technologies now make it possible to model biological systems more accurately than ever before [@tanay_scaling_2017]. One area where single-cell data has been particularly useful is in the study of cellular dynamic processes, such as the cell cycle, cell differentiation and cell activation [@etzrodt_quantitativesinglecellapproaches_2014]. Such dynamic processes can be computationally modelled using trajectory inference (TI) methods (also known as pseudemporal ordering methods), which use single-cell profiles from a population in which the cells are at different unknown points in the dynamic process  [@trapnell_definingcelltypes_2015; @cannoodt_computational_2016; @moon_manifoldlearningbasedmethods_2018]. These methods computationally order the cells along a trajectory topology, which can be linear, bifurcating, or more complex tree or graph structures. Because TI methods offer an unbiased and transcriptome-wide understanding of a dynamic process [@tanay_scaling_2017],  they allow the objective identification of new (primed) subsets of cells [@schlitzer_identificationcdc1cdc2committed_2015], delineation of a differentiation tree [@velten_humanhaematopoieticstem_2017; @see_mappinghumandc_2017] and inference of regulatory interaction responsible for one or more bifurcations [@aibar_scenicsinglecellregulatory_2017]. 
Current applications of TI focus on specific subsets of cells, but ongoing efforts to construct transcriptomic catalogues of whole organisms [@regev_scienceforumhuman_2017; @han_mappingmousecell_2018] underline the urgency for accurate, scalable[@aibar_scenicsinglecellregulatory_2017; @angerer_singlecellsmake_2017] and user-friendly TI methods.

```{r}
add_sfig(
  figure_file("methods_timeline.svg", "4-method_characterisation"),
  "methods_over_time", 
  "New TI methods over time. a) Number of methods published or in preprint. b) Number of methods which can handle a particular type of trajectory."
)
```

<!-- §2: Plethora of new TI methods -->
A plethora of TI methods has been developed over the last years, and even more are being created every month (`r ref("sfig", "methods_over_time", "a")`). We gathered a list of `r sum(implementations$is_ti, na.rm=TRUE)` TI methods (`r ref("table", "implementations")`), which we believe to be the most complete list to date. It is perhaps surprising that almost all of them have a unique combination of characteristics, in terms of the required inputs (prior information), produced outputs (topology fixing and trajectory type) and methodology used.

```{r, results='asis' }
add_table(
  "implementations",
  read_rds(figure_file("implementations_table.rds", "4-method_characterisation")),
  "Overview of the trajectory inference methods included in this study, and several characteristics thereof. This table will be continuously updated at https://goo.gl/FnsTxG."
)
```
<!-- §3: Topology inference -->
One distinctive characteristic of TI methods is whether the topology of the trajectory is inferred computationally, or was fixed by design.
Early TI methods typically fixed the topology algorithmically (e.g. linear [@bendall_singlecelltrajectorydetection_2014; @schlitzer_identificationcdc1cdc2committed_2015; @shin_singlecellrnaseqwaterfall_2015; @campbell_bayesiangaussianprocess_2015] or bifurcating [@haghverdi_diffusionpseudotimerobustly_2016; @setty_wishboneidentifiesbifurcating_2016]), or through parameters provided by the user [@trapnell_dynamicsregulatorscell_2014; @matsumoto_scoupprobabilisticmodel_2016]. These methods therefore mainly focused on correctly ordering the cells along this fixed topology. 
Other methods attempt to infer the topology computationally, which increases the difficulty of the problem at hand, but allows these methods  to be broadly applicable on more use cases. Methods that perform topology inference are still in the minority, though current trends suggest this will soon change (`r ref("sfig", "methods_over_time", "c")`).  A particularly interesting development is presented in the AGA method [@wolf_graphabstractionreconciles_2017] which is the only TI method currently able to deal with disconnected graphs.

<!-- §4: Prior information -->
Another key characteristic of TI methods is the selection of prior information that a method requires or can optionally exploit. Prior information can be supplied as  a starting cell from which the trajectory will originate, a set of important marker genes, or even a grouping of cells into cell states. Providing prior information to a TI method can be both a blessing and a curse. In one way, prior information can help the method to find the correct trajectory among many, equally likely, alternatives. On the other hand, incorrect or noisy prior information can bias the trajectory towards current knowledge. Moreover, prior information is not always easily available, and its subjectivity can therefore lead to multiple equally plausible solutions, restricting the applicability of such TI methods to well studied systems.

<!-- §5: Components -->
A reductionist approach to characterising TI methods consists in dissecting them into a set of algorithmic components, as any component can have a significant impact on the performance, scalability, and output data structures. Across all TI methods, these components can be broadly grouped into two stages; (i) conversion to a simplified representation using dimensionality reduction, clustering or graph building and (ii) ordering the cells along the simplified representation [@cannoodt_computational_2016]. Interestingly, components are frequently shared between different algorithms (`r ref("sfig", "implementation_components")`). For example, minimal spanning trees (MST), used in the first single-cell RNA-seq trajectory inference methods [@trapnell_dynamicsregulatorscell_2014], is shared by almost half of the methods we evaluated  (`r ref("sfig", "implementation_components", "b")`).

```{r}
add_sfig(
  figure_file("implementation_components.svg","4-method_characterisation"), 
  "implementation_components", 
  "TI methods frequently share steps."
)
```

<!-- §6: Problem statement -->
Given the  diversity in TI methods, an important issue to address is a quantitative assessment of the performance and robustness of the existing TI methods. Many attempts at tackling  this issue have already been made[@haghverdi_diffusionpseudotimerobustly_2016; @ji_tscanpseudotimereconstruction_2016; @welch_slicerinferringbranched_2016; @matsumoto_scoupprobabilisticmodel_2016; @duverle_celltreebioconductorpackage_2016; @cannoodt_scorpiusimprovestrajectory_2016; @lonnberg_singlecellrnaseqcomputational_2017; @campbell_probabilisticmodelingbifurcations_2017], but due to the high number of TI methods available today and the great diversity in the outputted data structures, a comprehensive benchmarking evaluation of TI methods is still lacking.
This is problematic, as new users to the field are confronted with a wide array of TI methods, without a clear idea about what method would be the most optimal for their problem. 
Moreover, the strengths and weaknesses of existing methods need to be assessed, so that new developments in the field can focus on improving the current state-of-the-art.

# Results
In this study, we performed a comprehensive evaluation for `r sum(methods$evaluated)` TI methods (Overview: `r ref("fig", "evaluation_overview", "a")`, Extended overview: `r ref("sfig", "extended_evaluation_overview")`). The inclusion criterion for TI methods was primarily based on their free availability, the presence of a programming interface, and the date of publication  (`r ref("table", "implementations")`). Only methods published before june 2017 are included in the current version of the evaluation, while more recent methods will be added in the next version. The evaluation comprised three core aspects: (i) source-code and literature-based characterisation of TI methods, (ii) assessment of the accuracy and scalability of TI methods by comparing predicted trajectories with a gold standard, and (iii) a quality control of the provided software and documentation.

In order to make gold standard trajectories and predicted trajectories directly comparable to one another, we developed a common probabilistic model for representing trajectories from all possible sources (`r ref("fig", "evaluation_overview", "b")`). In this model, the overall topology is represented by a network of "milestones", and the cells are placed within the space formed by each set of connected milestones. We defined a set of metrics for comparing the likeness of such trajectories, each assessing a different aspect of the trajectory: the similarity in cell ordering, the cell neighbourhood and the topology (`r ref("fig", "evaluation_overview", "c")`). The data compendium consisted of both synthetic datasets, which offer the most exact gold standard, and real datasets, which offer the highest biological relevance. These real datasets come from a variety of single-cell technologies, organisms, and dynamic processes, and contain several types of trajectory topologies (`r ref("sfig", "dataset_characterisation")` and `r ref("stable", "datasets")`). To generate synthetic datasets, we simulated a gene regulatory networks using a thermodynamic model [@schaffter_genenetweaversilicobenchmark_2011], and subsequently simulated a single cell profiling experiment of different single-cell RNA-seq technologies (`r ref("sfig", "synthetic_data_generation")`).

```{r}
add_sfig(
  figure_file('dataset_characterisation.svg', '2-dataset_characterisation'),
  "dataset_characterisation",  
  "Characteristics of the real datasets used for the evaluation."
)
```

```{r}
add_sfig(
  figure_file('synthetic_data_generation.svg', "2-dataset_characterisation/3-synthetic"),
  "synthetic_data_generation",
  "Synthetic data generation workflow."
)
```

As the aim of this study is to provide both guidelines for the use of TI methods and a base for the development of new TI methods, we not only assessed the accuracy of the predictions made by a trajectory, but also investigated the quality of the method's implementation. To do this, we scored each method using a checklist of important scientific and software development practices, including software packaging, documentation, automated code testing, and peer review. This allowed us to rank each method based on its user friendliness, developer friendliness, and potential broad applicability on new unseen datasets. Finally, using both the benchmark results and quality control, we  produced a flow chart with practical guidelines for selecting the most appropriate TI method for a given use case.

`r anchor("fig", "evaluation_overview")`
![](`r figure_file("ti_evaluation_overview_v3.svg", experiment_id="evaluation_overview")`)

`r ref("fig", "evaluation_overview")` **An overview of several key aspects of the evaluation.** a) . b) In order to make any two trajectories comparable with one another, a common trajectory model was used to represent gold standard trajectories from the real and synthetic datasets, as well as any predicted trajectories from TI methods. c) Three metrics were defined in order to assess the similarity in cell ordering, cell neighbourhood, and topology, for any two trajectories.

----

```{r}
add_sfig(
  figure_file('expanded_ti_evaluation_overview_v1.svg', 'evaluation_overview'),
  "extended_evaluation_overview",
  "Extended overview", 
  15, 
  7.5
)
```

## Evaluation of trajectory inference methods
<!-- §1: Introduce figure --> 
An overview of the main results from this study is shown in `r ref("fig", "results_overview")`. This includes an overview of the results obtained from the method characterisation (`r ref("fig", "results_overview", "a")`), the benchmarking evaluation (`r ref("fig", "results_overview", "b")`), and the quality control evaluation (`r ref("fig", "results_overview", "c")`).

`r anchor("fig", "results_overview")`
![](`r figure_file("overview.svg", experiment_id="9-main_figure")`)

`r ref("fig", "results_overview")` **Overview of the results on the evaluation of `r sum(methods$evaluated)` TI methods.** a) The methods were characterised according to the maximum complexity trajectory they can infer, whether or not the inferred topology is constrained by the algorithm or a parameter, and which prior information a method requires or can optionally make use of. b) The methods are ordered according to the overall score achieved in the benchmark evaluation. Also shown are the aggregated scores per metric, source and trajectory type, as well as the average execution time across all datasets and the percentage of executions in which no output was produced. c) Overall performance in the quality control evaluation is highly variable, even amongst the highest ranked methods according to the benchmark evaluation. Also listed are the quality control scores aggregated according to practicality and the different categories.

-------

<!-- §2: Overall and topology types -->
Having ordered all methods by their overall benchmarking score, we found that Slingshot predicted the most accurate trajectories, followed by TSCAN and Monocle DDRTree. When we looked at the benchmark scores per trajectory type, Slingshot was to be the only method that performs well across most trajectory types. However, we found that several methods were specialised in predicting specific trajectory types; for example SCORPIUS for linear trajectories, reCAT for cycles, Monocle DDRTree for rooted trees, and Wishbone for multifurcations. This last result is unexpected, as Wishbone is not able to output any multifurcating topologies. We will discuss the effect of topology types and constraints on inferrable trajectory types placed by the method in Section SEC_TRAJTYPE.

<!-- §3: real versus synthetic -->
We observed a high correlation (0.7-0.9) between results originating from real datasets versus those originating from synthetic datasets (`r ref("sfig", "real_synthetic_comparison")`). This confirms both the relevance of the synthetic data and the accuracy of the gold standard in the real datasets. However, this correlation was lower for converging and multifurcating datasets, potentially because only a small number of real datasets were available for such topologies (`r ref("stable", "datasets")`).

```{r}
add_sfig(
  figure_file(
    "real_synthetic_comparison.svg",
    "5-optimise_parameters/5-real_synthetic_comparison"
  ), 
  "real_synthetic_comparison", 
  "Comparison between performance on real and synthetic datasets across trajectory types."
)
```

<!-- §4: Compare metrics -->
As the different metrics were selected to assess the correctness of a trajectory in various approaches, it is expected to observe differences in rankings of TI methods amongst the different metrics. While we saw no direct link between the edge flip scores versus the correlation or RF MSE metric, methods that obtained a high correlation score also tended to obtain higher RF MSE scores (`r ref("sfig", "metric_comparison")`). In addition, the methods that were able to detect more complex trajectory types also obtained higher edge flip scores, in comparison to methods whose trajectory topology was fixed to a simple trajectory topology. We will explore this issue in more detail in a further section SEC_TRAJTYPE.

```{r}
add_sfig(
  figure_file('metrics_comparison.svg', "5-optimise_parameters/7-metric_comparison"), 
  "metric_comparison", 
  "Comparison between the different metrics across trajectory types."
)
```

During the execution of a method on a dataset, if the time limit (>6h) or memory limit (32GB) was exceeded, or an error was produced, a zero score was returned for that execution. If a method consistently generated errors on this dataset across all replicates, the error is called “data-specific”, otherwise “stochastic”. Several methods obtained a high overall score despite having 5 to 10% of failed executions (e.g. Monocle DDRTree, SLICE, Wishbone), meaning these methods could rank even higher if not for the failed executions. Methods that do not scale well with respect to the number of cells or genes will exceed the time or memory limits for the largest datasets (reCAT, SCOUP, StemID). For a few methods, the time or memory limits were exceeded too often, making the benchmarking results uncomparable to those of other methods (SCIMITAR, Ouija, Pseudogp).

### Trajectory types and topology complexity

In most cases, the methods which were specifically designed to  handle a particular trajectory type, also performed better on data containing this particular trajectory type (`r ref("sfig", "trajtype_handle_comparison")`). These methods had typically better edge flip scores - as can be expected - and RF MSE scores, compared to methods not able to handle the particular trajectory type. However, the correlation score typically followed the opposite pattern, where methods restricted to linear trajectory types, such as embeddr, SCORPIUS and phenopath, produced the best ordering, irrespective of whether the dataset contained a linear trajectory or more complex trajectories.
To further investigate the effect of trajectory complexity on performance, we divided them in two groups: linear methods (restricted to linear and cyclic topologies) and non-linear methods. While there were no significant differences in performance on linear datasets, non-linear methods had significantly higher edge flip scores but lower correlation scores on datasets containing more complex trajectory types  (`r ref("sfig", "linear_split_comparison")`). Together, this indicates that current non-linear methods potentially contain less accurate ordering algorithms. A combination of the ordering methods from the top linear methods, combined with the topology inference from top non-linear methods, could therefore be a possibility for future research.

```{r}
add_sfig(
  figure_file('trajtype_handle_comparison.svg', "8-compare_topology"), 
  "trajtype_handle_comparison", 
  "ADD CAPTION"
)
```

```{r}
add_sfig(
  figure_file('linear_split_comparison.svg', "8-compare_topology"), 
  "linear_split_comparison", 
  "Linear methods have a higher correlation score on non-linear datasets. Methods were split in linear methods (those which can either only detect cyclic or linear trajectories) and non-linear methods."
)
```

Despite their similar overall performance, the topologies predicted by the top methods differed considerably. Trajectories detected using the default parameters of slingshot and cellTree tended to be simpler, while those detected by TSCAN and Monocle DDRTree gravitated towards more complex topologies (`r ref("fig", "topology_complexity_comparison")`). Monocle DDRTree, for example frequently predicted a tree topology, even when only a cyclic, linear or bifurcating topology was present in the data (`r ref("fig", "topology_complexity_comparison", "a")`). Trajectories generated by Monocle DDRTree (at default parameters) tended to contain more nodes and edges (`r ref("fig", "topology_complexity_comparison", "b")`), which could give this method an advantage on datasets with complex tree topologies, but could also explain its relatively low performance on linear and bifurcating datasets. Indeed, when we assessed how often a method can infer the correct topology, slingshot and TSCAN were good at correctly predicting linear, bifurcating and converging trajectories, while Monocle DDRTree was by far the best method to infer tree topologies. Nonetheless, the overall accuracy of topology prediction was very low, with Slingshot correctly predicting bifurcating and converging topologies for half of the datasets, and Monocle DDRTree predicting the correct tree topology in 12% of the cases (`r ref("sfig", "topology_sensitivity")`). Inferring the correct topology without parameter tuning therefore still is an open challenge. Conversely, when the data contains a complex trajectory structure, TI will currently still require a considerable guidance by the end user, to either optimise the parameters or to choose the method with which the output best fits the user's expert knowledge.

`r anchor("fig", "topology_complexity_comparison")`
![](`r figure_file("topology_complexity_comparison.svg","8-compare_topology")`)

`r ref("fig", "topology_complexity_comparison")` **INSERT CAPTION.** Caption subdetails.

----

```{r}
add_sfig(
  figure_file('topology_sensitivity.svg', "8-compare_topology"), 
  "topology_sensitivity", 
  "ADD CAPTION"
)
```

### Effect of prior information

In the current version of the evaluation, we only provided prior information when a method required it . We did not observe a major difference in method performance between methods which did and did not receive prior information (`r ref("sfig", "required_priors_performance_comparison")`). Rather, methods which received prior information were on average positioned in the middle of the ranking.  Furthermore, we could not find any dataset where methods which received prior information performed significantly better than other methods.

```{r}
add_sfig(
  figure_file('required_priors_performance_comparison.svg', "10-prior_information"), 
  "required_priors_performance_comparison", 
  "Comparison of method performance."
)
```

### Algorithm components

We assessed whether the components of an algorithm can be predictive for method performance, using both random forest classification and statistical testing. Methods which include principal curves (such as Slingshot and SCORPIUS), k-means (which include Slingshot and several other top scoring methods) and some graph building (which include almost every top scoring method) tended to have a slightly higher performance (`r ref("sfig", "method_components_scores")` and `r ref("stable", "component_importance")`). On the other hand, methods using t-SNE and ICA for their dimensionality reduction were ranked lower, although this was not statistically significant.

```{r}
add_sfig(
  figure_file("method_components_scores.svg","6-developer_guidelines"), 
  "method_components_scores", 
  "Score of methods containing particular algorithmic components. NS: Not significant. *: "
)
```

```{r}
add_stable(
  read_rds(figure_file(
    "component_importance.rds",
    "6-developer_guidelines"
  ))[[params$table_format]], 
  "component_importance", 
  "Importance values for algorithmic components."
)
```

## Method quality control

While not directly related to the accuracy of the inferred trajectory, the quality of the implementation is also an important evaluation metric, since good unit testing makes sure the implementation is correct, good documentation makes it easier for potential users to apply the method on their data, and overall good code quality makes it possible for other developers to adapt the method and extend it further. We therefore looked at the implementation of each method, and assessed their quality using a transparent scoring scheme (`r ref("stable", "qc_checks")`). The individual quality checks can be grouped in two ways: what aspect of the method they investigate (availability, code quality, code assurance, documentation, method behaviour and the depth by which the method was presented in its study) or which purpose(s) they serve (user friendliness, developer friendliness or future proof). These categorisations can help current developers to improve their tool, and guide the selection of users and developers to use these tools for their purpose. After publishing this preprint, we will contact the authors of each method, allowing them to improve their method before the final publishing of the evaluation.

```{r}
add_stable(
  read_rds(figure_file(
    "qc_checks.rds",
    "4-method_characterisation"
  ))[[params$table_format]], 
  "qc_checks", 
  "Scoring scheme for method quality control."
)
```

We found that most methods fulfilled the basic criteria, such as free availability and basic code quality criteria (`r ref("fig", "implementation_qc_overview")`). We found that recent methods had a slightly better quality than older methods (`r ref("sfig", "qc_over_time")`). However, several quality aspects were consistently lacking for the majority of the methods (`r ref("fig", "implementation_qc_overview", " right")`) and we believe that these should receive extra attention from developers. Although these outstanding issues cover all five categories, code assurance and documentation in particular are problematic areas, notwithstanding several studies pinpointing these as good practices [@wilson_best_2014; @artaza_top10metrics_2016].

`r anchor("fig", "implementation_qc_overview")`
![](`r figure_file("implementation_qc_overview.svg","4-method_characterisation")`)

`r ref("fig", "implementation_qc_overview")` **Overview of the quality control results for every method.**

----

```{r}
add_sfig(
  figure_file("qc_over_time.svg","4-method_characterisation"), 
  "qc_over_time", 
  "Quality control scores and number of citations over time."
)
```
We observed no clear relation between method quality and method performance (`r ref("fig", "results_overview")` and `r ref("sfig", "qc_performance_comparison")`). We could also not find any quality aspect which was significantly associated with method performance.

```{r}
add_sfig(
  figure_file(
    "qc_performance_comparison.svg",
    "5-optimise_parameters/6-qc_performance_comparison"
  ), 
  "qc_performance_comparison", 
  "Comparison between method qc (x-axis) and method performance (y-axis)."
)
```

## Practical guidelines

Based on the results of our benchmark, we created a set of practical guidelines for method users `r ref("fig", "user_guidelines")`. As method's performance is heavily dependent on the trajectory type being studied, the choice of method will by primarily driven by the prior knowledge of the user about what trajectory topology is expected in the data. For the majority of use cases, the user will know very little about the expected trajectory, except perhaps whether the data is expected to contain multiple trajectories, cycles or a complex tree structure. In each such use case, a different set of methods performed optimally, with Monocle DDRTree performing best when the data contains a complex tree, and Slingshot performing better when the trajectory could as well be less complex. For the use case where the data is expected to contain multiple trajectories or cycles, no methods were included in the current version of the evaluation, although AGA`r cite_methods("aga")` for disconnected trajectories, and Waddington-OT`r cite_methods("wadot")` and AGA`r cite_methods("aga")` for regular graphs are currently the only methods in literature able to handle these types of trajectories. If the user knows the exact topology, our evaluation suggests reCAT for cycles, SCORPIUS for linear trajectories, and Slingshot for bifurcating trajectories, although in the last case other topologies could also be returned by the method if it better fits the data. The most difficult use case is when the topology is known but more complex than a bifurcating or cyclic trajectory. Here, to our knowledge, only ElPiGraph.R [github.com/Albluca/ElPiGraph.R](https://github.com/Albluca/ElPiGraph.R), which is not yet published, can be used.

`r anchor("fig", "user_guidelines")`
![](`r figure_file("tree.svg", experiment_id="7-user_guidelines")`)

`r ref("fig", "user_guidelines", anchor=TRUE)` **Practical guidelines for method users.**

----

# Discussion
<!-- The context -->
Trajectory inference is unique among most other categories of single-cell analysis methods, such as clustering, normalisation and differential expression, because it models the data in a way that was almost impossible using bulk data. Indeed, for no other single-cell analysis types have so many tools been developed, according to several repositories such as omictools.org [@henry_omictoolsinformativedirectory_2014], the "awesome single cell software" list [@davis_awesomesinglecelllistsoftware_2018] and scRNA-tools.org [@zappia_exploringsinglecellrnaseq_2017]. It is therefore critical that these methods, now reaching `r sum(methods$is_ti, na.rm=TRUE)`, are evaluated to guide the users in their choice. In this preprint, we summarise the initial version of our evaluation of these methods, focusing on the quality control and the accuracy of their model using the default parameters on real and synthetic data. Comparing the maximum overall score over time, it is encouraging to see multiple incremental improvements over the state-of-the-art (`r ref("sfig", "leading_method")`). We believe that the benchmarking presented in this study will pave the way to the next few state-of-the-art improvements.

```{r}
add_sfig(
  figure_file('leading_method.svg', "9-main_figure"), 
  "leading_method", 
  "Method performance compared to the time the method was published. The red line indicates the optimal performance up to a point in time."
)
```

<!-- Future work -->
In this study, we presented our first version of evaluation of TI methods. We are convinced that these results are already useful for both method users and tool developers, as we provide clear practical guidelines for users depending on their current knowledge of the trajectory, as well as an objective benchmark on which new methods can be tested. Nonetheless, our evaluation can be expanded on several points, all of which we will try to tackle in the near future:

* Inclusion of more recent methods, especially those which can detect more complex trajectory types such as general graphs `r cite_methods(c("wadot", "calista"))` and disconnected graphs `r cite_methods("aga")`.
* A parameter tuning for each method, both across all trajectory types, as well as on specific trajectory types.
* A test of robustness on noise, parameter changes and dataset size
* An evaluation of the stability of method's results when running the same parameter setting on the same dataset multiple times
* Evaluate the methods when given optional priors, and compare with performance without priors
* Assess the effect of noisy or incorrect prior information
* Test methods on datasets with no clear trajectory present
* Include feedback on the quality control scoring scheme, and update the scoring when method get updated
* Test the scalability of each method, both in the number of cells and the number of features
* Include more real datasets with complex trajectory types, as they become available
* Evaluate methods on other single-cell -omics datasets, such as proteomics and epigenomics data

<!-- Are certain methods better than others? -->
Our evaluation indicated a large difference between current TI methods, with Slingshot, followed by TSCAN and Monocle DDRTree, towering above all other methods. Nonetheless, we found that methods which did not perform well across all settings, can still be useful in certain specific use cases. Indeed, on data containing more than one bifurcations, Monocle DDRTree clearly performed better than other methods. We found that this was mainly because the default parameters of Monocle DDRTree prefer are more likely to detect tree topologies, while those of Slingshot prefer to find linear and bifurcation trajectories.

<!-- Our common output model, advantages & limitations -->
We managed to wrap the output of all methods into one common format. This not only allowed us to compare different methods with a gold standard, but could also be useful for TI users because it allows the user to test multiple methods without manual conversion of input and output. Furthermore, it makes it possible to compare the output of different methods, which opens up possibilities for new comparative visualisation techniques or ensemble methods. However, we acknowledge that our model has some limitations. It can currently not take into account uncertainty of the cell's position. This can be both locally, for example uncertainty of its position within one branch, or globally, for example uncertainty over connections between branches. Some current methods already model this uncertainty in some way, mainly locally `r cite_methods("ouija")`, and in the future we will adapt our output model to also allow this uncertainty.

<!-- Synthetic data: observations and possibilities -->
The use of synthetic data to evaluate methods can be problematic, because the model to generate the data is not necessarily similar to the way a true biological system works. This could bias an evaluation towards methods for which the underlying model best fits the model used for generating data. Therefore, it is essential that results of an evaluation on synthetic data are confirmed using real data. In our study, the overall performance of methods was very similar between real and synthetic datasets, confirming the biological relevance of the synthetic data. Given this, we believe our synthetic data can now be used to effectively prototype new TI methods, which use ever more complex models, such as disconnected graphs. Furthermore, it is expected that in the future methods will be able to model even more complex behavior, such as multiple dynamic processes happening at parallel in a single cell, the integration of datasets from different patients or trajectories in a spatial context [@tanay_scaling_2017; @cannoodt_computational_2016]. Synthetic data based on our workflow could therefore be used to spearhead the development of these methods, given that currently only a limited number of real datasets are available for which the methods could be useful. Furthermore, we believe our evaluation workflow could also be used to evaluate other types of single-cell modelling techniques, such as single-cell network inference, clustering and normalisation. We sincerely hope that such efforts will lead to a more rapid development of accurate methods, and will in the near future provide a package which can be used to simulate synthetic data for a wide variety of single-cell modelling problems.

# Methods

## Trajectory inference methods

### Method selection
We gathered a list of `r sum(methods$is_ti, na.rm=TRUE)` trajectory inference methods (`r ref("table", "implementations")`), by searching in literature for "trajectory inference" and "pseudotemporal ordering", and based on two existing lists found online [@davis_awesomesinglecelllistsoftware_2018; @gitter_singlecellpseudotimeoverviewalgorithms_2018].
A continuously updated list can also be found [online](https://goo.gl/FnsTxG)). We welcome any contributions by creating an issue at [github.com/dynverse/dynverse/issues](https://github.com/dynverse/dynalysis/issues).

Methods were excluded from the evaluation based on several criteria: `r glue::collapse(glue::glue("({non_inclusion_reasons$footnote}) {non_inclusion_reasons$long}"), ", ")`. This resulted in the inclusion of `r sum(methods$evaluated)` methods in the evaluation (`r ref("table", "implementations")`).

### Method input
As input, we provided for each method either the raw count data (after cell and gene filtering) or normalised expression values, based on the description in the methods documentation or from the study describing the method. Furthermore, when required, we also provided a maximum of 7 types of prior information. This prior information was extracted from the gold/silver standards as follows:

* **Start cells** The identity of one or more start cells. For both real and synthetic data, a cell was chosen which was the closest (in geodesic distance) from each milestone with only outgoing edges. For ties, one random cell was chosen. For cyclic datasets, a random cell was chosen.
* **End cells** The identity of one or more end cells. Similar as the start cells, but now for every state with only ingoing edges.
* **# end states** Number of terminal states. Number of milestones with only ingoing edges.
* **Grouping** For each cell a label to which state/cluster/branch it belongs. For real data, the states from the gold/silver standard. For synthetic data, each milestone was seen as one group, and cells were assigned to its closed milestone.
* **# branches** Number of branches/intermediate states. For real data, the number of states in the gold/silver standard. For synthetic data, the number of milestones.
* **Time course** For each cell a time point from which it was sampled. If available, directly extracted from the gold standard. For synthetic data: the simulation time.

### Common trajectory model
Due to the absence of a common format for trajectory models, most methods return a unique set of output formats with few overlap. We therefore created wrappers around each method (available at [github.com/dynverse/dynmethods](https://www.github.com/dynverse/dynmethods)) and postprocessed its output into a common probabilistic trajectory model (`r ref("sfig", "wrapping", "a")`). This model consisted of three parts. (i) The milestone network represents the overall network topology, and contains edges between different milestones and the length of the edge between them. (ii) The milestone percentages contain for each cell its position between milestones, and sums for each cell to one. (iii) The trajectory model also allows for regions of delayed commitment, where cells are allowed to be positioned between three or more connected milestones. Regions must be explicitly defined in the trajectory model. Per region, one milestone must be directly connected to all other milestones in the network.

```{r}
add_sfig(
  figure_file('wrapping.svg', '4-method_characterisation'),
  "wrapping", 
  "**A common interface for TI methods.** a) The input and output of each TI method is standardised. As input, each TI method receives either raw or normalised counts, several parameters, and a selection of prior information. After its execution, a method uses one of the five wrapper functions to transform its output to the common trajectory model. This common model then allows to perform common analysis functions on trajectory models produced by any TI method. b) The specific transformations performed by each of the wrapper functions is explained in more detail.", 
  15, 
  7.5
)
```

Depending on the output of a method, we used different strategies to convert the output to our model (`r ref("sfig", "wrapping", "b")`). Special conversions are denoted by an \*.

```{r}

# create functions to list the methods within a particular conversion category
list_sentence <- function(x) {
  pritt(glue::collapse(x[1:length(x) - 1],", "), " and ", last(x))
}
conversion <- function(conversion_oi) {
 methods_evaluated %>% 
   filter(map_lgl(conversion_split, ~ conversion_oi %in% .)) %>% 
   mutate(text = paste0(method_name, ifelse(conversion_special, "\\*", ""))) %>%
   pull(text) %>% 
   list_sentence()
}
```

* **Type 1, direct:** Methods which assign each cell to a branch together with a pseudotime value and a branch network: `r conversion("branching_local")`. The branch network is used as the milestone network, the percentages of a cell are proportional with its branch pseudotime.
* **Type 2, linear pseudotime:** Methods returning a single pseudotime value: `r conversion("linear")`. A single edge between two milestones, its percentage proportional to the pseudotime.
* **Type 3, end state probability:** Methods returning both a global pseudotime value and a probability for every end state: `r conversion("branching_global")`. We use a single start state and add an edge to every end milestone each representing an end state. The global pseudotime then determines the distance from the begin milestone, the rest of the cell's position is calculated by distributing the residual percentage over the end states, proportionally to the end state probabilities.
* **Type 4, cluster assignment:** Methods returning a cluster assignment and a cluster network:  `r conversion("branching_cluster")`. The cluster network was used as milestone network, each cell received percentage 1 or 0 based on its cluster assignment.
* **Type 5, projection onto nearest branch:** Methods returning a cluster assignment, cluster network and dimensionality reduction: `r conversion("branching_local_projection")`. We projected each cell on the closest point on the edges between its own cluster and neighbouring clusters within the dimensionality reduction. This projection allowed us to give each cell a pseudotime within the edge, which was then converted into our model as described above, using the cluster network as milestone network.

Special conversions were necessary for certain methods:

* **DPT** We projected the cells onto the cluster network, consisting of a central milestone (this cluster contains the cells which were assigned to the "unknown" branch) and three terminal milestones, each corresponding to a tip point. This was then processed as described above.
* **Sincell** To constrain the number of milestones this method creates, we merged two cell clusters iteratively until the percentage of leaf nodes was below a certain cutoff, default at 25%.  This was then processed as described above.
* **SLICER** Although this method returns a branch network and a global pseudotime, the method did not return a position of each cell on the individual branches. We therefore used the cell graph provided by the method to calculate the shortest paths between the start and end cells found by the method, and then used this graph as the backbone for all other cells, creating a local pseudotime for every cell within its branch. The output was then further processed as above.
* **SLICE** As discussed in the vignette of SLICE, we ran principal curves one by one for every edge detected by SLICE. This local pseudotime was then processed as above.
* **MFA** We used the branch assignment as state probabilities, which together with the global pseudotime were processed as described above.

## Real datasets
We gathered a list of real datasets by searching for "single-cell" at the Gene Expression Omnibus and selecting those datasets in which the cells are sampled from different stages in a dynamic process (`r ref("stable", "datasets")` and `r ref("sfig", "dataset_characterisation")`). The scripts to download and process these datasets will be made available n our repository ([github.com/dynverse/dynalysis](http://www.github.com/dynverse/dynalysis)). When available, we preferred to start from the raw counts data. These raw counts were all normalised and filtered using a common pipeline, discussed later. We determined a reference standard for every dataset using labelling provided by the author's, and classified the standards into gold and silver based on whether this labelling was determined by the expert using the expression (silver standard) or using other external information (such as FACS or the origin of the sample, gold standard) (`r ref("stable", "datasets")`).

```{r}
add_stable(
  read_rds(figure_file("datasets.rds","2-dataset_characterisation"))[[params$table_format]], 
  "datasets", 
  "Real datasets used in this study."
)
```

## Synthetic datasets

Our workflow to generate synthetic data is based on the well established workflow used in the evaluation of network inference methods [@schaffter_genenetweaversilicobenchmark_2011; @marbach_wisdom_2012] and consists of four main steps: network generation, simulation, gold standard extraction and simulation of the scRNA-seq experiment (`r ref("sfig", "synthetic_data_generation")`). At every step, we took great care was taking to mimic real cellular regulatory networks as best as possible, while keeping the model simple and easily extendable. For every synthetic dataset, we used a random real dataset as a reference dataset (from those described earlier), making sure the number of variable genes and cells were similar.

### Network generation

One of the main process involved in cellular dynamic processes is gene regulation, where regulatory cascades and feedback loops lead to progressive changes in expression and decision making. The exact way a cell chooses a certain path during its differentiation is still an active research field, although certain models have already emerged and been tested in vivo. One driver of bifurcations seem to be mutual antagonism, where genes [@xu_regulationbifurcatingcell_2015] strongly repress each other, forcing one of the two to become inactive [@graf_forcingcellschange_2009]. Such mutual antagonism can be modelled and simulated [@wang_quantifyingwaddingtonlandscape_2011; @ferrell_bistabilitybifurcationswaddington_2012]. Although such a two-gene model is simple and elegant, the reality is frequently more complex, with multiple genes (grouped into models) repressing each other [@yosef_dynamicregulatorynetwork_2013].

To simulate certain trajectory topologies, we therefore designed module networks in which the cells will follow a particular trajectory topology given certain parameters (`r ref("sfig", "modulenets")`). Two module networks generated linear trajectories (linear and linear long), two generated simple forks (bifurcating and converging), one generated a complex fork (trifurcating), one generated a rooted tree (consecutive bifurcating) and two generated trajectories which simple undirected graphs (bifurcating loop and bifurcating convergence).

From these module networks we generated gene regulatory networks in two steps: generation of the main regulatory network, and addition of extra target genes from real regulatory networks. For each dataset, we used the same number of genes as were differentially expressed in the real datasets. 5% of genes were assigned to be part of the main, and were randomly distributed among all modules (with at least one gene per module). We sampled edges between these individual genes (according to the module network) using an uniform distribution between 1 and the number of possible targets in each module. To add additional target genes to the network, we assigned every regulator from the network to a real regulator in a real network (from regulatory circuits [@marbach_tissuespecificregulatorycircuits_2016]), and extracted for every regulator a local network around it using personalized pagerank (with damping factor set to 0.1), as implemented in the `page_rank` function of the *igraph* package.

```{r}
add_sfig(
  read_rds(figure_file(
    "modulenets.rds",
     experiment_id = "2-dataset_characterisation/3-synthetic"
   )), 
  "modulenets", 
  "Module networks, milestone networks and the module dynamics for each of the different types of synthetic data", 
  15, 
  5
)
```

### Simulation of thermodynamics

To simulate the gene regulatory network, we used a system of differential equations similar to which has been standard in evaluations of network inference [@marbach_wisdom_2012]. In this model, changes in gene expression ($x_i$) and protein expression ($y_i$) are modeled using ordinary differential equations [@schaffter_genenetweaversilicobenchmark_2011] (ODEs):

$$
\begin{aligned}
\label{eq:mrna_ode}
\frac{dx_i}{dt} &= \underbrace{m \times f(y_1, y_2, ...)}_\text{production} - \underbrace{\lambda \times x_i}_\text{degradation}
\end{aligned}
$$
$$
\begin{aligned}
\label{eq:prot_ode}
\frac{dy_i}{dt} &= \underbrace{r \times x_i}_\text{production} - \underbrace{\Lambda \times y_i}_\text{degradation}
\end{aligned}
$$

where $m$, $\lambda$, $r$ and $\Lambda$ represent production and degradation rates, the ratio of which determines the maximal gene and protein expression. The two types of equations are coupled because the production of protein $y_i$ depends on the amount of gene expression $x_i$, which in turn depends on the amount of other proteins through the activation function $f(y_1, y_2, ...)$.

The activation function is inspired by a thermodynamic model of gene regulation, in which the promoter of a gene can be bound or unbound by a set of transcription factors, each representing a certain state of the promoter. Each state is linked with a relative activation $\alpha_j$, a number between 0 and 1 representing the activity of the promoter at this particular state. The production rate of the gene is calculated by combining the probabilities of the promoter being in each state with the relative activation:

$$
\begin{aligned}
\label{eq:input_function_probabilities}
f(y_1, y_2, ..., y_n) = \sum_{j \in \{0, 1, ..., n^2\}} \alpha_j \times P_j
\end{aligned}
$$

The probability of being in a state is based on the thermodynamics of transcription factor binding. When only one transcription factor is bound in a state:
$$
\begin{aligned}
P_j \propto \nu = \left(\frac{y}{k}\right)^{n}
\end{aligned}
$$

Where the hill coefficient $n$ represents the cooperativity of binding and $k$ the transcription factor concentration at half-maximal binding. When multiple regulators are bound:
$$
\begin{aligned}
P_j \propto \nu =  \rho \times \prod_j \left(\frac{y_j}{k_j}\right)^{n_j}
\end{aligned}
$$

where $\rho$ represents the cooperativity of binding between the different transcription factors.

$P_i$ is only proportional to $\nu$ because $\nu$ is normalized such that $\sum_{i} P_i = 1$.

To each differential equation, we added an additional stochastic term: 
$$
\begin{aligned}
\frac{dx_i}{dt} = m \times f(y_1, y_2, ...) - \lambda \times x_i &+ \eta \times \sqrt{x_i} \times \Delta W_t \\
\frac{dy_i}{dt} = r \times x_i - \Lambda \times y_i &+ \eta \times \sqrt{y_i} \times \Delta W_t
\end{aligned}
$$

with $\Delta W_t \sim \mathcal{N}(0, h)$.

Similar to [@schaffter_genenetweaversilicobenchmark_2011], we sample the different parameters from random distributions, given in `r ref("stable", "samplers")`.

```{r}
add_stable(
  read_rds(figure_file(
    "samplers.rds",
    "2-dataset_characterisation/3-synthetic")
  )[[params$table_format]], 
  "samplers", 
  "Distributions from which each parameter in the thermodynamic model was sampled."
)
```

These SDEs were simulated using the Euler–Maruyama approximation, with time-step $h = 0.01$ and noise strength $\eta = 8$. The total simulation time varied between 5 for linear and bifurcating datasets, 10 for consecutive bifurcating, trifurcating and converging datasets, 15 for bifurcating converging datasets and 30 for linear long, cycle and bifurcating loop datasets. The burn-in period was for each simulation 2. Each network was simulated 32 times.

### Simulation of the single-cell RNA-seq experiment

For each dataset we sampled the same number of cells as were present in the reference real dataset, limited to the simulation steps after burn-in. Next, we used the Splatter package [@zappia_splattersimulationsinglecell_2017] to estimate the different characteristics of a real dataset, such as the distributions of average gene expression, library sizes and dropout probabilities. We used Splatter to simulate the expression levels $\lambda_{i,j}$ of housekeeping genes $i$ (to match the number of genes in the reference dataset) in every cell $j$. These were combined with the expression levels of the genes simulated within a trajectory. Next, true counts were simulated using $Y'_{i,j} \sim \text{Poi}(\lambda_{i,j})$. Finally, we simulated dropouts by setting true counts to zero by sampling from a Bernoulli distribution using a dropout probability $\pi^D_{i,j} =\frac{1}{1+e^{-k(\text{ln}(\lambda_{i,j})-x_0)}}$.

### Gold standard extraction

Because each cellular simulations follows the trajectory it its own speed, knowing the exact position of a cell within the trajectory topology is not straightforward. Furthermore, the speed by which simulated cells made a decision between two or more alternative paths was highly variable. To estimate a cell's position during a simulation within the trajectory topology, we therefore used the known progression of the modules, given in `r ref("sfig", "modulenets", "c")`,  as a backbone. We smoothed the expression in each simulation using a rolling mean with a window of 50 time steps, and then calculated the average module expression along the simulation. We used dynamic time warping, implemented in the dtw R package [@giorgino_computingvisualizingdynamic_2009; @tormene_matchingincompletetime_2009], with an open end to align a simulation to all possible module progressions, and then picked the alignment which minimised the normalised distance between the simulation and the backbone. In case of cyclical trajectory topologies, the number of possible milestones a backbone could progress through was limited to 20.

## Common normalisation pipeline
We used a standard single-cell RNA-seq preprocessing pipeline which applies parts of the scran and scater Bioconductor packages [@lun_stepbystepworkflowlowlevel_2016]. The advantages of this pipeline is that it works both with and without spike-ins, and includes a harsh cell filtering which looks at abnormalities in library sizes, mitochondrial gene expression, and number of genes expressed using median absolute deviations (set to 3). We required that a gene has to be expressed in at least 5% of the cells, and that it should have an average expression higher than 0.05. Furthermore, we used the pipeline to select the most highly variable genes, using a false discovery rate of 5% and a biological component higher than 0.5. As a final filter, we removed both all zero genes and cells until convergence.

## Evaluation metrics
We used three different metrics to assess the performance of each TI method (`r ref("fig", "evaluation_overview")`). The importance of using multiple metrics to compare complex models has been stated repeatedly [@norel_selfassessmenttrap_2011]. Indeed, each of our metrics looks at the correspondence between prediction and gold standard in a different way: the correlation between geodesic distances assesses the overall ordering, the RF prediction error assesses the cell neighbourhood, and the edge flip score assesses the milestone network topology (`r ref("sfig", "score_aspects")`).

```{r}
add_sfig({
  read_rds(figure_file(
    "score_aspects.rds", 
    experiment_id="3-score_rules/5-score_aspects"))
  }, 
  "score_aspects", 
  "Different scores assess different aspects of the correspondence between a prediction and gold standard trajectory.",
  15, 
  8
)
```

### Correlation between geodesic distances
The accuracy of cell ordering was assessed by first calculating the geodesic distances $d(c_i, c_j)$ between all cells in both the gold standard and the prediction. The geodesic distance is calculated by finding the shortest path between two cells through the milestone network, while taking into account the edge lengths as returned by most methods. To calculate this for every pair of cells, we first calculated the shortest path lengths $d(m_a, m_b)$ between all milestones using Dijkstra's algorithm, as implemented in the R *igraph* package. We also calculated $d(m_a, c_i)$ for every cell to its neighbouring milestones by calculating the earth's movers distance using the shortlist method [@gottschlich_shortlistmethodfast_2014], as implemented in the R *transport* package. Then we again used Dijkstra's algorithm to calculate the shortest path lengths between every pair of cells which minimises:

$$
d(c_i, c_j) = min_{a \in A, b \in B} d(m_a, c_i) + d(m_a, m_b) + d(m_b, c_j)
$$

where A and B contain all milestones adjacent to respectively $c_i$ and $c_j$.
The final score was then calculated the Spearman's rank correlation between all $d(c_i, c_j)$ in gold standard and prediction.

In some cases, methods are unable to place some cells on a trajectory, mainly because they classify these cells as noisy. In these cases, we add these cells to a new milestone and put this milestone at a distance of 5 times the maximal distance between two milestones within the network.

### Random forest prediction error

Although the correlation between geodesic distances directly assesses the position of the cells in the trajectory, a bad correlation does not directly imply that similar cells were not grouped together by the method, as illustrated in `r ref("sfig", "score_aspects", c("b", "c"))`. For example, certain methods will inevitably reach an incorrect ordering because they cannot handle the correct trajectory type, but these methods could still correctly place similar cells next to each other. We therefore also included a metric which looks at the local neighbourhood of each cell, and assesses whether this neighbourhood can accurately predict the position of this cell in the gold standard. We used a Random Forest regression, implemented in the ranger package [@wright_rangerfastimplementation_2017] to separately predict milestone percentages of every cell in the gold standard, using the milestone percentages of these cells in the prediction as features. We then used the out-of-bag mean-squared error on these percentages to score each method's capability of predicting the correct neighbourhood of each cell.

### Edge flip score

As a third independent score, we assessed the similarity between the milestone network topology. We first simplified each network, by merging consecutive linear edges into one edge, by adding new milestones within self loops such that $A \rightarrow A$ would be converted $A \rightarrow B \rightarrow C \rightarrow D$, by adding an intermediate node to linear networks. Because we are interested , the network was made undirected. Next, we define the edge flip score as the minimal number of edges which should be added or removed to convert one network into the other, divided by the total number of edges in both networks. This problem is equivalent to the maximum common edge subgraph problem, a known NP-hard problem with no scalable solutions [@bahiense_maximumcommonedge_2012]. We implemented a branch & bound approach for this problem, by first enumerating all possible edge additions and removals with the minimal number of edges (the edge difference between the two networks) and if none of the new networks was isomorphic, try out all solutions with an additional two edge changes. To further limit the search space, we make sure the degree distributions between the two networks are similar, before assessing whether the two networks are isomorphic using the BLISS algorithm [@junttila_engineeringefficientcanonical_2007], as implemented in the R *igraph* package.

```{r}
add_sfig(
  figure_file("edge_flip_comparison.svg", experiment_id="3-score_rules/4-edge_flip"), 
  "edge_flip_comparison", 
  paste0("Edge flip score values for common trajectory topologies, defined in ", ref("sfig", "modulenets")), 
  15, 
  5
)
```

### Aggregation of scores
```{r}
add_sfig(
  figure_file("metric_aggregation_overview_v1.svg", experiment_id="evaluation_overview"), 
  "aggregation_of_results", 
  "Aggregation methodology of metric scores.", 
  15, 
  5
)
```

## Benchmark

### Pipeline
<!-- Look at dyneval methods to write this paragraph -->

### Method execution
We ran each method in a controlled environment, where we limited the memory usage (max. 32GB) and total run time (max 24h). Moreover, we disabled the function of the R `set.seed` function, to avoid that methods false make their output deterministic.

### Effect of prior information

To assess the effect of prior information on the performance of a method, we compared the performance of methods which do or do not require the prior information using a two-tailed Mann–Whitney U test. P-values were controlled for multiple testing using Benjamini-Hochberg correction.

### Effect of algorithm components
To assess the effect of algorithm components on the performance of a method, we used (1) a two-tailed Mann–Whitney U test, as implemented in the `wilcox.test` R function, and (2) increase in node purity importance scores using random forest classification, as implemented in the R randomForest package, predicting whether a method scored better than the median score. To evaluate whether a method increased or decreased performance, we used the estimate of the location parameter of the Mann–Whitney U test. We only investigated algorithm components which were part of at least 4 different methods in our evaluation study. P-values were controlled for multiple testing using Benjamini-Hochberg correction.

## Method quality control

```{r}
qc_references <- readRDS(derived_file("checks.rds", "4-method_characterisation")) %>% 
  .$references %>% 
  str_split("[, \n]+") %>% 
  unlist() %>% 
  unique() %>% 
  discard(is.na) %>%
  glue::collapse(., ";") %>%
  paste0("[", ., "]")
```

We created a transparent scoring scheme to check the quality of each method (`r ref("stable", "qc_checks")`), based on several existing tool quality and programming guidelines in literature and online `r qc_references` . The goal of this quality control in the first place is to stimulate the improvement of current methods, and the development of user and developer friendly new methods. The quality control assessed 6 categories, each looking at several aspects, which are further divided into individual items. The availability category checks whether the method is easily available, whether the code and dependencies can be easily installed, and how the method can be used. The code quality assesses the quality of the code both from a user perspective (function naming, dummy proofing and availability of plotting functions) and a developer perspective (consistent style and code duplication). The code assurance category is frequently overlooked, and checks for code testing, continuous integration [@beaulieu-jones_reproducibility_2017] and an active support system. The documentation category checks the quality of the documentation, both externally (tutorials and function documentation) and internally (inline documentation). The behaviour category assesses the ease by which the method can be run, by looking for unexpected output files and messages, prior information and how easy the trajectory model can be extracted from the output. Finally, we also assessed the certain aspects of the study in which the method was proposed, such as publication in a peer-reviewed journal, the number of dataset on which the usefulness of the method was shown, and the scope of method evaluation in the paper.

Each aspect was further assigned to one or more applications, based on whether it influenced the user friendliness of the tool (such as code availability, good documentation or contains plotting functions), the developer friendliness of the tool (such as unit testing, inline documentation and a clear licensing), or indications that the tool will be broadly applicable on new datasets (such as being open-source, containing a good tutorial and support system, and being thoroughly evaluated in the study where it was published).

Each quality aspect received a weight by how frequently it was found in several papers and online sources which discuss tool quality (`r ref("table", "implementations")`). This was to make sure that more important aspect, such as the open source availability of the method, outweigh other aspects, such as the availability of a graphical user interface. Within each aspect, we also assigned a weight to the individual questions being investigated (`r ref("table", "implementations")`). For calculating the final score, we weighed each of the six categories equally.

## Trajectory types

We classified all possible trajectory topologies into distinct trajectory types, based on topological criteria. These trajectory types start from the most general trajectory type, a disconnected directed graph, and moves down (within a directed acyclic graph structure), progressively becoming more simple until the two basic types: linear and cycles (`r ref("sfig", "trajectory_type_tree", "a")`). For every directed trajectory type, a corresponding undirected trajectory type can also be defined (`r ref("sfig", "trajectory_type_tree", "b")`). In most cases, a method which can detect a complex trajectory type, was also be able to detect less complex trajectory types, with the only exception being DPT (limited to bifurcating trajectories).

```{r}
add_sfig(
  figure_file("trajectory_type_trees.svg","2-dataset_characterisation/2-trajectory_types"), 
  "trajectory_type_tree", 
  "Different trajectory types can be classified in a directed acyclic graph structure, in which certain types are a generalisation of other trajectory types.", 
  15, 
  10
)
```

# Acknowledgements

# Supplementary Material

```{r "supplementary", echo=FALSE, results='asis'}
source("analysis/paper/sfigs_stables.R")
```

# Colophon

This report was generated on `r Sys.time()` using `r devtools:::platform_info()$version` and the following packages:
```{r colophon, cache = FALSE}
devtools::session_info()$packages %>% knitr::kable()
```

The current Git commit details are:

```{r}
git2r::repository(".")
```
I would avoid making "absolute" statements like this one.
Add the timecourse prior to the figure
Changed the tenses here to past (results usually only include past tense, except when referring to figures or later in the text)
why surprisingly?
because wishbone is only able to detect bifurcations. Should mention this in the text :)
Added an explanation in the following sentence
But still more should be added, I think this simply indicates that none of the methods are performing well on multifurcating datasets. But we should show this, eg. looking at the raw performance on multifurcating datasets versus other datasets. If we don't have the time to show this, then we should not mention wishbone here
LINK TO

### Trajectory types and topology complexity
LINK TO

### Trajectory types and topology complexity
ADD CAPTION
TODO
ADD CAPTION
FILL IN SIGNIFICANCE LEVELS
? Unclear what you mean by this.
It's already a bit better, though could be improved still!
to do: add metric figure
needs to be updated with waypoints
this is not used anymore
TODO
More to add here?
* Look at dynmethods::execute_method to write this paragraph
Are there more exectuion details?