### 2018-03-22 analysis/paper/supp_notes/supp_note_trajectory_inference_methods.Rmd
add supp note for acos

---
title: "A comparison of single-cell trajectory inference methods: towards more accurate and robust tools"
bibliography: ../references.bib
csl: ../nature-biotechnology.csl
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage[table]{xcolor}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{tabu}
  - \usepackage[normalem]{ulem}
  - \usepackage{lscape}
  - \usepackage{pbox}
  - \usepackage[labelformat=empty]{caption}
  - \ExecuteBibliographyOptions{sorting=none,sortcites=true,autocite=superscript,autopunct=false}
mainfont: Open Sans
output: 
  pdf_document: 
    latex_engine: xelatex
    dev: png
    keep_tex: true
    citation_package: biblatex
---

*`r format(Sys.time(), '%d %B, %Y')`*

```{r setup, include=FALSE}
# knitr options
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.path = paste0(rprojroot::find_rstudio_root_file(), "/analysis/paper/supp_notes/.scratch/")
)

# load libraries
library(dynalysis)
library(dynplot)
library(tidyverse)
library(cowplot)
```
Wouter Saelens* ¹ ², Robrecht Cannoodt* ¹ ² ³, Helena Todorov¹ ² ⁴, Yvan Saeys¹ ²

\* Equal contribution  
¹ Data mining and Modelling for Biomedicine, VIB Center for Inflammation Research, Ghent, Belgium.  
² Department of Applied Mathematics, Computer Science and Statistics, Ghent University, Ghent, Belgium.  
³ Center for Medical Genetics, Ghent University Hospital, Ghent, Belgium.  
⁴ Centre International de Recherche en Infectiologie, Inserm, U1111, Université Claude Bernard Lyon 1, CNRS, UMR5308, École Normale Supérieure de Lyon, Univ Lyon, F-69007, Lyon, France

# Supplementary Note: Trajectory Inference Wrapppers

## `acos`: Arc cosine
This is a control method for inferring cyclic trajectories in a very simple approach,
consisting of a dimensionality reduction step and the computation of the arc cosine.

### Parameters
There is only one parameter, namely the dimensionality reduction method used.
```{r}
par_set <- description_acos()$par_set
par_set
```

The default parameters are the following:
```{r}
defprm <- ParamHelpers::dfRowToList(ParamHelpers::generateDesignOfDefaults(par_set, trafo = TRUE), par_set, 1)
defprm
```

### Example dataset
We will be using a cyclic toy dataset to demonstrate this method with.

```{r}
toy <- dyntoy::toy_tasks %>% filter(model == "cycle") %>% extract_row_to_list(row_id = 2)
expression <- toy$expression

plot_default(toy)
```

### Wrapper
In the first step, a dimensionality reduction to two dimensions is computed. By default, this is PCA. 
Here, it is assumed that the dimensionality reduction will be centered around the origin 0.

```{r}
dimred <- dynmethods:::dimred(expression, method = defprm$dimred, ndim = 2)

# Plot dimred
dimred_df <- data.frame(dimred) %>% rownames_to_column("cell_id")
ggplot(dimred_df) + geom_point(aes(Comp1, Comp2)) + coord_equal()

```

The ordering of the cells is computed by calculating the bearing of each cell with respect to the origin. The bearing is transformed to a [0,1] range.

```{r}
pseudotimes <- atan2(dimred[,2], dimred[,1]) / 2 / pi + .5

dimred_df <- dimred_df %>% mutate(pseudotime = pseudotimes[cell_id])
ggplot(dimred_df) + geom_point(aes(Comp1, Comp2, colour = pseudotime)) + coord_equal() + viridis::scale_color_viridis()
```

The prediction is wrapped using the `add_cycling_trajectory_to_wrapper` function. 
```{r}
prediction <- wrap_prediction_model(
  cell_ids = rownames(expression)
) %>% add_cyclic_trajectory_to_wrapper(
  pseudotimes = pseudotimes,
  do_scale_minmax = FALSE
) %>% add_dimred_to_wrapper(
  dimred = dimred
)

plot_default(prediction)
```

The figure below shows a comparison of the orderings in the toy versus the prediction.
```{r}
plot_strip_connections(toy, prediction)
```

# Colophon
This report was generated on `r Sys.time()` using `r devtools:::platform_info()$version` and the following packages:
```{r colophon, cache = FALSE}
devtools::session_info()$packages %>% knitr::kable()
```