### 2018-06-27 scripts/evaluate_toy/1-methods-toys.Rmd
reorganise dynalysis

---
title: "1-methods-toys"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Testing methods on simple toy datasets

```{r}
library(tidyverse)
library(dyntoy)
library(dyneval)
library(dynmethods)

set.seed(1)

data("toy_tasks")
tasks <- toy_tasks
```

```{r}
tasks$counts <- map(tasks$counts, function(counts) {
  cbind2(counts, matrix(sample(1:100, nrow(counts)*100, replace=TRUE), nrow=nrow(counts))) %>% {magrittr::set_colnames(., paste0("G", 1:ncol(.)))}
})
task <- dynutils::extract_row_to_list(tasks, 3)
pheatmap::pheatmap(task$counts)

source("~/thesis/projects/dynverse/dynmodular/dimred_wrappers.R")
# space <- task$counts %>% dimred_dp(2, neigen=2) %>% as.data.frame
space <- task$counts %>% dimred_mds(2) %>% as.data.frame
# space <- task$counts %>% dimred_lle(2) %>% as.data.frame
space <- space %>% bind_cols(task$prior_information$grouping_assignment %>% slice(match(rownames(task$counts), cell_id)))
space %>%  ggplot() + geom_point(aes(Comp1, Comp2, color=group_id))
```

We choose certain parameters for which we know that the method will find (or focus on finding) a linear trajectory, and will do this quickly
```{r}
method_params <- yaml::yaml.load_file("toy_params.yml")$linear
method_names <- names(method_params)
# method_names <- c("ouija", "slice")

method_ti_methods <- map(method_names, ~get(paste0("ti_", .))()) %>% set_names(method_names)

metric_names <- c("mean_R_nx", "auc_R_nx", "Q_local", "Q_global", "correlation")

method_names <- c("scorpius")
```

```{r}
# results <- qsub::qsub_lapply(names(method_ti_methods), function(method_name) {
results <- purrr::map(method_names, function(method_name) {
  library(dyneval)

  cat("Processing ", method_name, "\n", sep="")
  method <- method_ti_methods[[method_name]]
  params <- method_params[[method_name]]

  factory <- function(fun) {
    function(...) {
      warn <- err <- NULL
      res <- withCallingHandlers(
        tryCatch(fun(...), error=function(e) {
          err <<- conditionMessage(e)
          NULL
        }), warning=function(w) {
          warn <<- append(warn, conditionMessage(w))
          invokeRestart("muffleWarning")
        })
      list(res, warn=warn, err=err, method_name=method_name)
    }
  }

  # factory(function() {
    dyneval:::evaluate_ti_method(tasks, method, params, metrics = metric_names, timeout = 100)
  # }
  # )()
})
# }, qsub_environment = list2env(lst(method_ti_methods, tasks, metric_names, method_params)), qsub_config = qsub::override_qsub_config(memory = "10G", execute_before=c("module load python/x86_64/3.5.1"), wait=F))
```

Gather the output of the different methods, and check whether there were any errors
```{r}
scores <- purrr::map(results, function(result) {
  bind_cols(
    tibble(
      model = attr(result, "extras")$.models
    ),
    attr(result, "extras")$.summary
  )
}) %>% bind_rows()

# scores %>% saveRDS("results/toy_method_scores.rds")
```

```{r}
scores %>% filter(errored)

print(scores$error)
```

```{r fig.height=10, fig.width=10}
scores %>% left_join(tasks %>% select(id, trajectory_type), by=c("task_id"="id")) %>% 
  dplyr::select(-starts_with("time")) %>%
  gather(score_id, score, -method_name, -method_short_name, -method_id, -task_id, -error, -warning, -model, -trajectory_type) %>%
  filter(score_id %in% c("errored", "correlation", "robbie_network_score", "mean_R_nx")) %>% 
  ggplot() + geom_bar(aes(method_name, score, fill=score_id), stat="identity") + 
  facet_grid(trajectory_type~score_id) + coord_flip()
```