### 2018-08-08 manuscript/supplementary_notes/metric_conformity/assessment.Rmd
add score conformity note

---
title: "assessment"
output: html_document
params:
  assessment: NULL
---

```{r}
rule <- dynutils::extract_row_to_list(rules, which(rules$id == assessment$rule_id))
```

## `r knitr::include_graphics(rule$image_location)` `r label_long(assessment$rule_id)`

`r rule$description`

```{r}
width <- ifelse(is.null(assessment$plot_datasets$width), 1, assessment$plot_datasets$width)
height <- ifelse(is.null(assessment$plot_datasets$height), 1, assessment$plot_datasets$height)
```
```{r, fig.width=width, fig.height=height}
assessment$plot_datasets
```

```{r}
width <- ifelse(is.null(assessment$plot_scores$width), 1, assessment$plot_scores$width)
height <- ifelse(is.null(assessment$plot_scores$height), 1, assessment$plot_scores$height)
```
```{r, fig.width=width, fig.height=height}
assessment$plot_scores
```