### 2018-08-08 manuscript/supplementary_notes/metric_conformity/metric_conformity.Rmd

---
title: "Method conformity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(dynbenchmark)

experiment("03-metric_characterisation/01-metric_conformity")
```

```{r}
assessments <- read_rds(derived_file("assessments.rds"))
rules <- read_rds(derived_file("rules.rds"))

rules <- rules %>% 
  mutate(
    image_location = map_chr(id, function(id) figure_file(paste0("images/", id, ".png"))),
    image_found = file.exists(image_location)
  )

if (any(!rules$image_found)) {stop("Files not found: ", basename(rules$image_location[!rules$image_found]))}
```

```{r results='asis'}
walkdf(assessments, function(assessment) {
  cat(knitr::knit_child("assessment.Rmd", quiet=TRUE, envir = environment()))
  cat("\n------------------------------\n")
})
```