### 2018-08-09 scripts/01-datasets/02-synthetic/README.Rmd
add dataset readme

---
output: github_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synthetic datasets