### 2018-08-10 scripts/01-datasets/02-synthetic/README.Rmd
add some readme to the datasets

---
output: github_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synthetic datasets

Each synthetic dataset is based on some characteristics of some real datasets. These characteristics include:

* The number of cells and features
* The number of features which are differentially expressed in the trajectory
* Estimates of the distribution of the library sizes, average expression, dropout probabilities, ... estimated by [Splatter](https://github.com/Oshlack/splatter)

These are estimated in [01-estimate_platform.R](01-estimate_platform.R).

Next, we simulate datasets using different simulators:

* [Dyngen](https://github.com/dynverse/dyngen), simulations of regulatory networks which will produce a particular trajectory
* [PROSSTT](https://github.com/soedinglab/prosstt), simulations of tree topologies using random walks
* [Splatter](https://github.com/Oshlack/splatter), simulations of non-linear paths between different states
* [Dynplot](https://github.com/dynverse/dynplot), simulations of toy data using random expression gradients in a reduced space