### 2018-08-10 scripts/04-method_testing/README.Rmd

---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Method testing

Here, we test the methods on some small real and toy datasets, to capture errors early on and make sure all methods run fine on these "simple" datasets.

These testing results are in the end pushed to [https://dynverse.github.io/dynmethods_automated_testing/index.html](https://dynverse.github.io/dynmethods_automated_testing/index.html)