### 2018-08-13 manuscript/supplementary_notes/metrics/assessment.Rmd
update to metric characterisation

---
title: "assessment"
params:
  assessment: NULL
---

```{r}
rule <- dynutils::extract_row_to_list(rules, which(rules$id == assessment$rule_id))
```

\subsection{`r rule$name`}

\begin{center}
```{r, out.width="2in", out.height="10in", dpi = 300}
knitr::include_graphics(rule$image_location)
```

\end{center}

`r rule$description`

A metric conforms to this rule if: $`r rule$conforms_if`$

```{r}
assessment$conformity %>% 
  select(metric_id, conforms) %>% 
  arrange(metric_id) %>% 
  mutate(
    metric_id = label_metrics(metric_id, label_type = "latex") %>% paste0("$", ., "$") %>% forcats::fct_inorder(),
    conforms = kableExtra::cell_spec(
      ifelse(conforms, "\U2714", "\U2716"), 
      background = ifelse(conforms, "#2ECC40", "#FF4136"),
      format = "latex"
    )
  ) %>% 
  spread(metric_id, conforms) %>% 
  knitr::kable(format = "latex", escape = FALSE) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::row_spec(0, hline_after = FALSE)
```

`r rule$observation`

```{r}
width <- ifelse(is.null(assessment$plot_datasets$width), 10, assessment$plot_datasets$width)
height <- ifelse(is.null(assessment$plot_datasets$height), 10, assessment$plot_datasets$height)
```
```{r A, fig.width=width, fig.height=height, out.width=paste0(width/2, "in")}
assessment$plot_datasets
```

```{r}
width <- ifelse(is.null(assessment$plot_scores$width), 10, assessment$plot_scores$width)
height <- ifelse(is.null(assessment$plot_scores$height), 10, assessment$plot_scores$height)
```
```{r B, fig.width=width, fig.height=height, out.width=paste0(width/2, "in")}
assessment$plot_scores
```

\pagebreak