### 2018-08-29 scripts/03-methods/README.Rmd
update readmes

---
output: github_document
---

```{r include = FALSE}
library(tidyverse)
library(dynbenchmark)

knitr::opts_chunk$set(echo = FALSE)
```

# Trajectory inference methods

Here we

- Compile all the information we have about TI methods
- Group methods into sets of "tools"
- Download and process the information from the QC spread sheet (https://docs.google.com/spreadsheets/d/1Mug0yz8BebzWt8cmEW306ie645SBh_tDHwjVw4OFhlE)
- Create tables and figures to interpret the methods and quality control

```{r}
dynbenchmark::render_scripts_documentation()
```