### 2018-08-29 scripts/02-metrics/README.Rmd

---
output: 
  github_document:
    pandoc_args: --atx-headers
---

# Metrics for comparing two trajectories

Good metrics are crucial for an unbiased and comprehensive benchmarking. Here we test several possible metrics for comparing two trajectories.

```{r include=FALSE}
library(dynbenchmark)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
```

```{r, results = "asis"}
knit_child_readme("01-metric_characterisation")
```

```{r, results = "asis"}
knit_child_readme("02-metric_conformity")
```