### 2018-08-29 scripts/01-datasets/01-real/README.Rmd

---
output: github_document
---

# Real datasets

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The generation of the real datasets is divided in two parts:

```{r}
dynbenchmark::render_scripts_documentation()
```