### 2018-08-29 scripts/01-datasets/02-synthetic/README.Rmd

---
output: github_document
---

# Synthetic datasets

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Each synthetic dataset is based on some characteristics of some real datasets. These characteristics include:

* The number of cells and features
* The number of features which are differentially expressed in the trajectory
* Estimates of the distribution of the library sizes, average expression, dropout probabilities, ... estimated by [Splatter](https://github.com/Oshlack/splatter).

Here we estimate the parameters of these "platforms" and use them to simulate datasets using different simulators. Each simulation script first creates a design dataframe, which links particular platforms, different topologies, seeds and other parameters specific for a simulator.

The data is then simulated using wrappers around the simulators (see [/package/R/simulators.R](/package/R/simulators.R)), so that they all return datasets in a format consistent with dynwrap.

```{r}
dynbenchmark::render_scripts_documentation()
```