### 2018-08-29 scripts/05-scaling/README.Rmd

---
output: github_document
---

```{r include = FALSE}
library(tidyverse)
library(dynbenchmark)

knitr::opts_chunk$set(echo = FALSE)
```

# Scalability of methods

Here we test how well a method scales with increasing number of features (genes) and/or cells.

Each method is run on randomly generated data with increasing gene set and cell set sizes, and the execution times and memory usages are modelled using linear models.

```{r}
dynbenchmark::render_scripts_documentation()
```