### 2018-08-29 scripts/02-metrics/01-metric_characterisation/README.Rmd
add ability to link to results

---
output: 
  github_document:
    pandoc_args: --atx-headers
---

# Metric characterisation

```{r include=FALSE}
library(dynbenchmark)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
```

A first characterisation of the metrics. For each metric we:

- generate some examples to get some intuition on how the metric works
- test the robustness to a metric to stochasticity or parameters when appropriate

```{r, results = "asis"}
render_scripts_documentation()
```

The results of this experiment are available [here](`r link_to_results()`)