### 2018-08-30 scripts/02-metrics/README.Rmd

---
output: dynbenchmark::github_markdown_nested
---

# Metrics for comparing two trajectories

Good metrics are crucial for an unbiased and comprehensive benchmarking. Here we test several possible metrics for comparing two trajectories.

```{r include=FALSE}
library(dynbenchmark)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
```

```{r, results = "asis"}
knit_nest("01-metric_characterisation")
```

```{r, results = "asis"}
knit_nest("02-metric_conformity")
```