### 2018-08-30 scripts/README.Rmd

---
output: dynbenchmark::github_markdown_nested
---

```{r}
library(tidyverse)
library(dynbenchmark)
```

# Scripts

The scripts folder contains all the scripts necessary to fully reproduce the benchmarking manuscript in chronologically ordered subfolders. Each subfolder contains a readme file with further explanations of the different scripts and what they do:

```{r}
dynbenchmark::render_scripts_documentation()
```