### 2018-08-30 scripts/03-methods/README.Rmd

---
output: dynbenchmark::github_markdown_nested
---

# Trajectory inference methods

Here we

- Compile all the information we have about TI methods
- Group methods into sets of "tools"
- Download and process the information from the QC spread sheet (https://docs.google.com/spreadsheets/d/1Mug0yz8BebzWt8cmEW306ie645SBh_tDHwjVw4OFhlE)
- Create tables and figures to interpret the methods and quality control

```{r}
dynbenchmark::render_scripts_documentation()
```