### 2018-08-30 scripts/03-methods/README.Rmd
cleanup methods

---
output: dynbenchmark::github_markdown_nested
---

# Trajectory inference methods

Here we

- Compile all the information we have about TI methods
- Characterise the methods with regards to user and developer friendliness (method quality control)
- Characterise the methods with regards to prior information, underlying algorithm, possible detectable trajectory types, ...

```{r}
dynbenchmark::render_scripts_documentation()
```

```{r, results = "asis"}
dynbenchmark::knit_nest("01-gather_methods_information")
```