### 2018-08-30 scripts/03-methods/01-gather_methods_information/README.Rmd

---
output: dynbenchmark::github_markdown_nested
---

# Gathering all the information we have about the methods

Most information of the methods are contained within their respective containers (see the [dynmethods](https://github.com/dynverse/dynmethods) repository, [https://github.com/dynverse/dynmethods](https://github.com/dynverse/dynmethods)). We gather additional information from our google sheets (https://docs.google.com/spreadsheets/d/1Mug0yz8BebzWt8cmEW306ie645SBh_tDHwjVw4OFhlE), which also contains the quality control for each methods.

```{r}
dynbenchmark::render_scripts_documentation()
```