### 2018-08-31 scripts/02-metrics/01-metric_characterisation/README.Rmd
fix completely sensible error for when dynbenchmark is not imported in readmes

---
output: dynbenchmark::github_markdown_nested
---

# Metric characterisation

A first characterisation of the metrics. For each metric we:

- generate some examples to get some intuition on how the metric works
- test the robustness to a metric to stochasticity or parameters when appropriate

```{r}
dynbenchmark::render_scripts_documentation()
```

The results of this experiment are available [here](`r dynbenchmark::link_to_results()`)