### 2018-09-17 scripts/04-method_testing/README.Rmd
update method testing scripts readme

---
output: dynbenchmark::github_markdown_nested
---

# Method testing

Each method is tested on a few small real and toy datasets. Ideally, each method should be able to execute successfully on each of these datasets.
If a method is not able to infer a trajectory on any of the datasets, it will not be included in any of the next experiments.

Overview experiment:
```{r}
dynbenchmark::render_scripts_documentation()
```