### 2018-09-18 scripts/02-metrics/README.Rmd

---
output: dynbenchmark::github_markdown_nested
---

# Metrics for comparing two trajectories

Good metrics are crucial for an unbiased and comprehensive benchmarking. Here we test several possible metrics for comparing two trajectories.

```{r, results = "asis"}
dynbenchmark::knit_nest("01-metric_characterisation")
```

```{r, results = "asis"}
dynbenchmark::knit_nest("02-metric_conformity")
```