### 2018-09-21 manuscript/supplementary.Rmd
Update supplementary RMD

---
title: "Supplementary material for: \nA comparison of single-cell trajectory inference methods: towards more accurate and robust tools"
output: 
  dynbenchmark::pdf_manuscript:
      render_change: FALSE
---

```{r setup, include=FALSE}
library(dynbenchmark)
library(tidyverse)
```

```{r}
supplementary <- read_rds("supplementary.rds")
sfigs <- supplementary$sfigs
stables <- supplementary$stables
refs <- supplementary$refs
```

# Supplementary figures
```{r, results="asis"}
sfigs %>% pmap(dynbenchmark:::plot_fig, ref_type = "sfig") %>% walk(cat)
```

# Supplementary tables

```{r, results="asis"}
stables %>% pmap(dynbenchmark:::show_table, ref_type = "stable") %>% walk(cat)
```

```{r setup, include=FALSE}
# source(paste0(dynbenchmark::get_dynbenchmark_folder(), "/analysis/paper/setup.R"))
# 
# # knitr options
# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
# knitr::opts_chunk$set(
#   collapse = TRUE,
#   warning = FALSE,
#   message = FALSE,
#   echo = FALSE,
#   comment = "#>",
#   fig.path = paste0(rprojroot::find_rstudio_root_file(), "/analysis/paper/.scratch/")
# )
# 
# list2env(read_rds(derived_file("sfigs_stables.rds", "paper")), environment())
```

```{r "supplementary", echo=FALSE, results='asis'}
# source(paste0(dynbenchmark::get_dynbenchmark_folder(), "/analysis/paper/sfigs_stables.R"))
```