### 2018-09-29 manuscript/paper.Rmd
update paper

---
title: "A comparison of single-cell trajectory inference methods: towards more accurate and robust tools"
output: dynbenchmark::pdf_manuscript
---

```{r setup, include=FALSE}
library(dynbenchmark)
library(tidyverse)

tools <- read_rds(result_file("tools.rds", "03-methods"))
methods <- read_rds(result_file("methods.rds", "03-methods"))
```

Wouter Saelens* ¹ ², Robrecht Cannoodt* ¹ ² ³, Helena Todorov¹ ² ⁴, Yvan Saeys¹ ²

\* Equal contribution  
¹ Data mining and Modelling for Biomedicine, VIB Center for Inflammation Research, Ghent, Belgium.  
² Department of Applied Mathematics, Computer Science and Statistics, Ghent University, Ghent, Belgium.  
³ Center for Medical Genetics, Ghent University Hospital, Ghent, Belgium.  
⁴ Centre International de Recherche en Infectiologie, Inserm, U1111, Université Claude Bernard Lyon 1, CNRS, UMR5308, École Normale Supérieure de Lyon, Univ Lyon, F-69007, Lyon, France

*`r format(Sys.time(), '%d %B, %Y')`*

`www.github.com/dynverse/dynbenchmark/commit/`r git2r::revparse_single(".","HEAD")$sha``
```{r}
n_tools <- nrow(tools)
n_tools_evaluated <- nrow(filter(tools, evaluated))
n_methods_evaluated <- nrow(filter(methods, evaluated, source %in% c("tool", "adaptation", "offtheshelf")))
n_control_methods_evaluated <- nrow(filter(methods, evaluated, source %in% c("control")))
n_datasets_real <- nrow(list_datasets("real"))
n_datasets_synthetic <- nrow(list_datasets("synthetic"))
```
# Abstract

Recent technological advances allow unbiased investigation of cellular dynamic processes, by generating -omics profiles of thousands of single cells and computationally ordering these cells along a trajectory. Since 2014, at least →`r n_tools`← tools for trajectory inference have been developed, but due to high variability in their inputs and outputs, these methods are difficult to compare. As a result, a comprehensive assessment of the performance of trajectory inference methods is still lacking, and more importantly also guidelines for users in the field.

In this study, we evaluated a total of →`r n_methods_evaluated`← methods in terms of cellular →ordering, topology, scalability, and user friendliness←. Our results indicate that existing methods are typically better at ordering the cells rather than inferring the correct trajectory topology, and that there is no "one-fits-all" method. Instead, the choice toof method →mostly← depends on the trajectory topology present in the data. →To aid users with selecting the most appropriate method, we therefore developed a set of guidelines, available as an interactive app at `r print_url("guidelines.dynverse.org")`.← While the field certainly has matured significantly since its inception, further progress is still necessary in order to cope with increasingly complex and novel use cases. We hope to spearhead such developments and the field as a whole, by making our evaluation pipeline easily extensible and freely available at `r print_url("https://www.github.com/dynverse/dynverse")`.

# Introduction
<!-- §1: TI overview -->
Single-cell -omics technologies now make it possible to model biological systems more accurately than ever before [@tanay_scaling_2017]. One area where single-cell data has been particularly useful is in the study of cellular dynamic processes, such as the cell cycle, cell differentiation and cell activation [@etzrodt_quantitativesinglecellapproaches_2014]. Such dynamic processes can be modelled computationally using trajectory inference (TI) methods, which order cells along a trajectory by their overall similarity in -omics datasets, mainly transcriptomics [@trapnell_definingcelltypes_2015; @cannoodt_computational_2016; @moon_manifoldlearningbasedmethods_2018]. The resulting trajectories are most often linear, bifurcating or tree-shaped, but more recent methods also allow to identify more complex trajectory topologies such as cyclic [@liu_reconstructingcellcycle_2017] or disconnected graphs [@wolf_graphabstractionreconciles_2017]. TI methods offer an unbiased and transcriptome-wide understanding of a dynamic process [@tanay_scaling_2017],  thereby allowing the objective identification of new (primed) subsets of cells [@schlitzer_identificationcdc1cdc2committed_2015], delineation of a differentiation tree [@velten_humanhaematopoieticstem_2017; @see_mappinghumandc_2017] and inference of regulatory interactions responsible for one or more bifurcations [@aibar_scenicsinglecellregulatory_2017]. 
Current applications of TI focus on specific subsets of cells, but ongoing efforts to construct transcriptomic catalogues of whole organisms [@regev_scienceforumhuman_2017; @han_mappingmousecell_2018] underline the urgency for accurate, scalable[@aibar_scenicsinglecellregulatory_2017; @angerer_singlecellsmake_2017] and user-friendly TI methods.

<!-- §2: Plethora of new TI methods -->
A plethora of TI methods has been developed over the last years, and even more are being created every month (`r ref("stable", "tools")`). →Indeed, in several repositories listing single-cell tools,  `r print_url("www.omictools.org")` [@henry_omictoolsinformativedirectory_2014], the "awesome-single-cell" list (`r print_url("https://github.com/seandavi/awesome-single-cell")`) [@davis_awesomesinglecelllistsoftware_2018] and `r print_url("www.scRNA-tools.org")` [@zappia_exploringsinglecellrnaseq_2017], TI methods are one of the largest categories.← All methods have their own unique set of characteristics in terms of underlying algorithm, the prior information they require and the outputs they produce. Two of the most distinctive differences between TI methods are →whether they fix the topology of the trajectory, and what type(s) of graph topologies can be detected←. Early TI methods typically fixed the topology algorithmically (such as linear [@bendall_singlecelltrajectorydetection_2014; @schlitzer_identificationcdc1cdc2committed_2015; @shin_singlecellrnaseqwaterfall_2015; @campbell_bayesiangaussianprocess_2015] or bifurcating [@haghverdi_diffusionpseudotimerobustly_2016; @setty_wishboneidentifiesbifurcating_2016]), or through parameters provided by the user [@trapnell_dynamics_2014; @matsumoto_scoupprobabilisticmodel_2016]. These methods therefore mainly focused on correctly ordering the cells along the fixed topology. →Most recent methods also infer the topology [@street_slingshotcelllineage_2017; @wolf_graphabstractionreconciles_2017], which increases the difficulty of the problem at hand, but allows an unbiased identification of both the ordering inside a branch and the topology connecting these branches.←

```{r}
add_stable(
 result_file("tools.xlsx", "03-methods"),
  "tools",
paste0("→Overview of available trajectory inference tools, and whether they were included in this study.←"),
  ""
)
```

<!-- §3: Problem statement -->
Given the diversity in TI methods, an important issue to address is a quantitative assessment of their performance and robustness. Many attempts at tackling this issue have already been made [@haghverdi_diffusionpseudotimerobustly_2016; @ji_tscanpseudotimereconstruction_2016; @welch_slicerinferringbranched_2016; @matsumoto_scoupprobabilisticmodel_2016; @duverle_celltreebioconductorpackage_2016; @cannoodt_scorpiusimprovestrajectory_2016; @lonnberg_singlecellrnaseqcomputational_2017; @campbell_probabilisticmodelingbifurcations_2017], but a comprehensive benchmarking evaluation of TI methods, which compares a large set of methods on a large set of datasets, is still lacking. This is problematic, as new users to the field are confronted with an overwhelming choice of TI methods, without a clear idea about which method would optimally solve their problem. Moreover, the strengths and weaknesses of existing methods need to be assessed, so that new developments in the field can focus on improving the current state-of-the-art.

```{r}
add_fig(
  fig_path = raw_file("overview_evaluation_v6.svg", ""),
  ref_id = "ti_evaluation_overview", 
  caption_main = "Overview of several key aspects of the evaluation.", 
  caption_text = "a) A schematic overview of our evaluation pipeline. b) In order to make the trajectories comparable to each other, a common trajectory model was used to represent →reference trajectories← from the real and synthetic datasets, as well as any predicted trajectories from TI methods. c) Trajectories are automatically classified into one of nine trajectory types, with increasing complexity. d) →We defined four metrics each assessing the similarity of a different aspect of the trajectory: the topology, the branch assignment, the cell positions and the important features.←",
   width = 13,
  height = 8, 
)
```

```{r}
add_sfig(
  fig_path = raw_file("overview_evaluation_expanded_v1.svg", ""),
  ref_id = "overview_evaluation_expanded",
  caption_main = "Extended overview of our evaluation pipeline.", 
  caption_text = "From literature we extracted a quality control checklist, a list of trajectory inference methods, a set of real datasets containing a trajectory and real regulatory networks used to generate the synthetic data. We created a wrapper of each method, so that its output was transformed into a common probabilistic trajectory model, and used these to infer trajectories on all real and synthetic datasets using their default parameters. Using several similarity metrics, we compared the →reference trajectory← of the real and synthetic datasets with the inferred trajectories. We also evaluated the quality of each method using the quality control checklist, and used both the benchmark results and quality control results to produce a final set of guidelines for method's users.", 
  width = 15, 
  height = 7.5
)
```

→

# Results
## Method overview
<!-- Method characterisation -->
In this study, we performed a comprehensive evaluation of  `r n_methods_evaluated` TI methods (`r ref("fig", "ti_evaluation_overview", "a")`, `r ref("sfig", "overview_evaluation_expanded")`). We included methods primarily based on their free availability and presence of a programming interface (`r ref("stable", "tools")`). In order to make the outputs from different methods directly comparable to each other, we developed a common probabilistic model for representing trajectories from all possible sources (`r ref("fig", "ti_evaluation_overview", "b")`). In this model, the overall topology is represented by a network of "milestones", and the cells are placed within the space formed by each set of connected milestones. The output of each method was converted to this model using different wrapper types (`r ref("fig", "results_summary", "a")`).

When required, we also provided prior information to the method, which ranges from "soft" priors, such as start or end cells, to "harder" priors such as the known grouping of the cells (`r ref("fig", "results_summary", "a")`).

The main difference between trajectory inference methods is whether the method fixes the topology, and if it is free, what kind of topologies it can detect. We defined nine possible types of topology, ranging from the very basic (linear, cyclical and bifurcating) to the most complex (connected and disconnected graphs). Most methods either focus on inferring linear trajectories, or limit the search to tree or less complex topologies, with only a selected few attempting to infer cyclic or disconnected topologies (`r ref("fig", "results_summary", "a")`).

<!-- Overall evaluation results -->
Each method was evaluated on four core aspects: (i) benchmarking given a gold or silver standard on `r n_datasets_real` real and `r n_datasets_synthetic` synthetic datasets, (ii) their scalability with both the number of cells and features (iii) the stability of the predictions across multiple bootstrap replicates, and (iv) the user friendliness of the provided software and documentation. Overall, we found a large diversity across all four evaluation criteria, with only a few methods, such as PAGA, GNG, Slingshot, Projected TSCAN and SCORPIUS, performing well across the board  (`r ref("fig", "results_summary", "b")`). We will first discuss each evaluation criterion in detail, after which we conclude with guidelines for method users and future perspectives for method developers.
```{r}
add_fig(
  fig_path = result_file("results_summary.pdf", experiment_id = "08-summary"),
  ref_id = "results_summary", 
  caption_main = glue::glue("A characterisation of the {n_methods_evaluated} methods evaluated in this study, and their overall evaluation results."), 
  caption_text = "(a) We characterised the methods according to their required priors, the types of inferreable topologies, whether or not the inferred topology is constrained by the algorithm or a parameter. The methods are grouped vertically based on the most complex trajectory type they can infer. (b) Overall results of the evaluation on four criteria: the benchmarking using a reference trajectory on real and synthetic data, the scalability with increasing number of cells and features, the stability across dataset bootstraps, and the quality control of the tool which implemented the method. ",
  width = 15,
  height = 18
)
```

```{r}
add_fig(
  fig_path = result_file("results_detailed.pdf", experiment_id = "08-summary"),
  ref_id = "results_detailed", 
  caption_main = glue::glue("Detailed results of the four main evaluation criteria: benchmarking, scalability, stability and quality control."), 
  caption_text = "(a) Results of the benchmark across metrics, dataset sources and dataset trajectory types. The performance of a method is generally more stable across dataset sources, but very variable depending on the metric and trajectory type. Also shown is how often a method errored, and what the reasons were for these errors. (b) Scalability results, with increasing number of cells and features. Shown are predicted times at a fixed set of cells (10.000) or features (10.000). (c) Stability results (d) Quality control results across six categories.",
  width = 15,
  height = 18
)
```
## Benchmark
<!-- Benchmark methods -->
We defined several possible metrics to compare a prediction to a reference trajectory (`r ref("snote", "metrics")`). Based on an analysis of their robustness and conformity to a set of rules (`r ref("snote", "metrics")`), we chose 4 metrics each assessing a different aspect of a trajectory: the topology (`r label_metric("him")`), the quality of clustering of the branch assignment  (`r label_metric("F1_branches")`), the cell positions (`r label_metric("correlation")`) and the quality of differentially expressed features  (`r label_metric("featureimp_wcor")`)(`r ref("fig", "ti_evaluation_overview", "d")`).The data compendium consisted of both synthetic datasets, which offer the most exact reference trajectory, and real datasets, which offer the highest biological relevance. These real datasets come from a variety of single-cell technologies, organisms, and dynamic processes, and contain several types of trajectory topologies (`r ref("stable", "datasets")`).  Some real datasets were classified as gold if the reference trajectory was defined using other data than the expression data, such as via cellular sorting. All other real datasets were classified as silver. For synthetic datasets we used several data simulators, including our own which simulates gene regulatory networks using a thermodynamic model of gene regulation [@schaffter_genenetweaversilicobenchmark_2011]. For each simulation, we used a real dataset as a reference, to match its dimensions, number of differentially expressed genes, drop-out rates and other statistical properties.

<!-- Benchmark overview -->
We found that method performance was very variable across datasets, indicating that there is no "one-size-fits-all" method that works well on every dataset `r ref("sfig", "benchmark_interpretation", "a")`. Even methods which can detect most of the trajectory types, such as PAGA or GNG, were not the best methods across all trajectory types `r ref("fig", "results_detailed", "a")`. The average performance between the different dataset sources was moderately to highly similar correlated, with the performance on all dataset sources being moderately to highly correlated (0.5 - 0.9) with the one on real gold datasets `r ref("sfig", "benchmark_interpretation", "b")`. On the other hand, the different metrics frequently disagreed with each other, with Monocle scoring better on the topology scores, while other methods, such as Slingshot, were better at ordering the cells and placing them into the correct branches `r ref("fig", "results_detailed", "a")`.

<!-- Topologies-->
The performance of a method was strongly dependent on the type of trajectory present in the data `r ref("fig", "results_detailed", "a")`. Slingshot typically performed better on datasets containing more simple topologies, while PAGA, pCreode and RaceID/StemID had higher scores on datasets with trees or more complex trajectories. This was reflected in the types of topologies detected by every method, as those predicted by Slingshot tended to contain less branches, while those detected by PAGA, pCreode and Monocle DDRTree gravitated towards more complex topologies (`r ref("sfig", "benchmark_interpretation", "c")`).

<!-- Complementarity between methods -->
The high variability between datasets, together with the diversity in detected topologies between methods, could indicate some complementarity between the different methods. To test this, we looked for subsets of methods at which the probability of getting a "top" model (defined as having a score at least 95% as the best model) was highest. On data which contains a tree trajectory, the use of the best method, RaceID / StemID, would result in getting top model about 50% of the time. This probability rapidly increased to up to 90% with the addition of 2 to 5 other methods (`r ref("fig", "complementarity", "a")`). We found similar indications of complementarity between the top methods on data containing only linear, bifurcation or multifurcating trajectories (`r ref("fig", "complementarity", "b")`). When the datasets from different trajectory types were pooled, the number of methods necessary to achieve more than 90% top models increased to more than 8. This pooling also resulted in a more diverse set of methods, containing both strictly linear methods such as SCORPIUS, and methods with a broad trajectory type range such as PAGA. Altogether, this shows that there is considerably complementarity between the different methods, and that users should try out different methods on their data.

```{r echo=FALSE, results="asis"}
add_fig(
  fig_path = result_file("complementarity.rds", experiment_id = "10-benchmark_interpretation"),
  ref_id = "complementarity", 
  caption_main = "Complementarity between different TI methods", 
  caption_text = "",
  width = 12,
  height = 6
)
```

```{r echo=FALSE, results="asis"}
add_sfig(
  fig_path = result_file("benchmark_interpretation.rds", experiment_id = "10-benchmark_interpretation"),
  ref_id = "benchmark_interpretation", 
  caption_main = "Robustness of the benchmark", 
  caption_text = "Shown are the distributions of the difference in size between predicted and reference topologies. A positive difference means that the topology predicted by the method is more complex than the one in the reference.",
  width = 12,
  height = 15
)
```

## Scalability
Given the ever increasing 
## Stability
←
## Quality control

While not directly related to the accuracy of the inferred trajectory, the quality of the implementation is also an important evaluation metric [@_doesyourcode_2018]. To assess this, we scored each method using a transparent checklist of important scientific and software development practices, including software packaging, documentation, automated code testing, and publication into a peer-reviewed journal (`r ref("stable", "qc_scorsheet")`).

We found that most methods fulfilled the basic criteria, such as free availability and basic code quality criteria (`r ref("fig", "results_detailed", "c")` and `r ref("sfig", "qc_overview")`). While recent methods had a slightly better quality score than older methods, several quality aspects were consistently lacking for the majority of the methods (`r ref("sfig", "qc_overview", " right")`) and we believe that these should receive extra attention from developers. Although these outstanding issues cover all five categories, code assurance and documentation in particular are problematic areas, notwithstanding several studies pinpointing these as good practices [@wilson_best_2014; @artaza_top10metrics_2016].

```{r}
add_sfig(
  fig_path = result_file("qc_overview.rds", experiment_id = "03-methods/02-tool_qc"),
  ref_id = "qc_overview", 
  caption_main = "Overview of the quality control results for every method.", 
  caption_text = glue::glue("Shown is the score given for each method on every item from our quality control score sheet ({ref('stable', 'qc_scoresheet')}). Each aspect of the quality control was part of a category, and each category was weighted so that it contributed equally to the final quality score. Within each category, each aspect also received a weight depending on how often it was mentioned in a set of papers discussing good practices in tool development and evaluation. This is represented in the plot as the height on the y-axis. Top: Average QC score for each method. Right: The average score of each quality control item. Shown into more detail are those items which had an average score lower than 0.5."),
   width = 15,
   height = 18
)
```

```{r}
add_stable(
 result_file("qc_scoresheet.xlsx", "03-methods/02-tool_qc"),
  "qc_scoresheet",
paste0("→Scoring sheet for method quality control. Each quality aspect was given a weight based on how many times it was mentioned in a set of articles discussing best practices for tool development.←"),
  ""
)
```

We observed no clear relation between method quality and method performance (`r ref("fig", "results_summary")`). We could also not find any quality aspect which was significantly associated with method performance.

# Discussion

<!-- The context -->
In this study, we presented a large scale evaluation of the performance of  `r n_methods_evaluated`  TI methods. By using a common structure and four metrics to compare the methods’ outputs, we were able to assess the performance of the methods on more than a two-hundred datasets, which allowed us to highlight the strengths and weaknesses of each method on →different types of trajectories. We also assessed several other important quality measures, such as quality of the method's implementation, the scalability to hundreds of thousands of cells, and the stability of the output on the same datasets.←

<!-- FOR METHOD USERS -->
<!-- Practical guidelines -->
Based on the results of our benchmark, we propose a set of practical guidelines for method users (`r ref("fig", "user_guidelines")` and `r print_url("guidelines.dynverse.org")`). We postulate that, as a method's performance is heavily dependent on the trajectory type being studied, the choice of method should be primarily driven by the prior knowledge of the user about which trajectory topology is expected in the data. For the majority of use cases, the user will know very little about the expected trajectory, except perhaps whether the data is expected to contain multiple disconnected trajectories, cycles or a complex tree structure. In each of these use cases, →our evaluation suggests a different set of optimal methods, as shown in `r ref("fig", "user_guidelines")`. Several other factors will also impact the choice of methods, such as the dimensions of the dataset, and the prior information which is available. These factors and several others can all be changed in our interactive app (`r print_url("http://dynverse.org")`). This app can also be used to explore the results of this evaluation, such as filtering the datasets, or changing the importance of the evaluation metrics for the final ranking.←

```{r}
add_fig(
  fig_path = result_file("guidelines.svg", experiment_id = "09-guidelines"),
  ref_id = "user_guidelines", 
  caption_main = "Practical guidelines for method users.", 
  caption_text = "As the performance of a method most heavily depends on the topology of the trajectory, the choice of TI method will be primarily influenced by the user's existing knowledge about the expected topology in the data. We therefore devised a set of practical guidelines, which combines the method's performance, user friendliness and the number of assumptions a user is willing to make about the topology of the trajectory. Methods to the right are ranked according to their performance on a particular (set of) trajectory type. Further to the right are shown the user friendliness scores (++: ≥ 0.95, +: ≥ 0.85, ± ≥ 0.75, - ≥ 0.65), overall performance (++: top method, +: difference between top method's performance ≥ -0.05, ±: ≥-0.2, -: ≥ -0.5) and required prior information.",
  width = 13,
  height = 10
)
```

<!-- Further considerations -->
When choosing a method, it is important to take two further points into account. First, it is critical that a trajectory, and the downstream results and/or hypotheses originating from it, are confirmed by multiple TI methods. This is to make sure the prediction is not biased due to the given parameter setting or the particular algorithms underlying a TI method. Second, even if the expected topology is known, it can be beneficial to also try out methods which make less assumptions about the trajectory topology. When the expected topology is confirmed using such a method, it provides extra evidence to the user's knowledge of the biological system. When a more complex topology is produced, →this could indicate that the underlying biology is much more complex than expected by the user.←

<!-- A common trajectory analysis framework -->
Critical to the broad applicability of TI methods is standardisation of the input and output interfaces of TI methods, so that users can effortlessly execute TI methods on their dataset of interest, compare different predicted trajectories, and apply downstream analyses such as find genes important for the trajectory, network inference [@aibar_scenicsinglecellregulatory_2017] or finding modules of genes [@saelens_comprehensiveevaluationmodule_2018]. Our framework is an initial attempt at tackling this problem, and we illustrate its usefulness here by comparing the predicted trajectories of several top-performing methods →on some a linear, a tree, a cyclic and a disconnected graph dataset (`r ref("fig", "example_predictions")`). Using our framework, this figure can be recreated by using only a couple of lines of R code (`r print_url("https://www.github.com/dynverse/dynmethods")`). This framework should still be extended to allow more input data, such as spatial and RNA velocity information [@manno_rnavelocitysingle_2018], and easier downstream analyses. In addition, further discussion within the field is required in order to arrive at a consensus concerning a common interface for trajectory models, which can include additional features such as uncertainty and gene importance.←

```{r}
add_fig(
  fig_path = result_file("example_predictions.rds", experiment_id = "11-example_predictions"),
  ref_id = "example_predictions", 
  caption_main = "Demonstration of how a common framework for TI methods facilitates broad applicability using an example dataset.", 
  caption_text = "",
  width = 13,
  height = 8
)
```

<!-- FOR METHOD DEVELOPERS -->
<!-- New tools: challenges and possibilities -->
Our study indicates that the field of trajectory inference is maturing, primarily for linear and bifurcating trajectories (as we illustrate with two datasets in `r ref("fig", "example_predictions", "a and b")`). Nonetheless, we also highlight several ongoing challenges, which should be addressed before TI can be a reliable and fast tool for analysing single-cell -omics datasets with complex trajectories. Foremost, new methods should focus on improving →the unbiased inference of tree, cyclic graph and disconnected topologies, as we found that methods repeatedly fail both on real and synthetic data, even if the trajectory can be easily identified in a dimensionality reduction (illustrated with some synthetic datasets in `r ref("fig", "example_predictions", "c and d")`). ← Furthermore, higher standards for code assurance, documentation and scalability could help the adoption of these tools across single-cell -omics field. To support the development of these new tools, we provide all datasets from this study [@cannoodt_goldstandardsinglecell_] and provide the code to compare methods in a series of R packages available at [github.com/dynverse/dynverse](https://github.com/dynverse/dynverse).

<!-- The robustness of the evaluation & development of new methods -->
We found that the performance of a method can be very variable between datasets, and therefore included a large set of both real and synthetic data within our evaluation, leading to a robust overall ranking of the different methods. Nonetheless, we also found that it is possible to define for almost every method a relatively large subset of datasets on which this method on average performs the best. These findings raise some important issues, as we want to avoid that a race to be the first on the ranking would stifle creativity, given that there is a high risk that a certain new methodology would not improve upon the state-of-the-art. Indeed, we firmly believe that "good-yet-not-the-best" methods [@norel_selfassessmenttrap_2011] can still provide a very valuable contribution to the field, especially if they make use of novel algorithms, return a more scalable solution, or provide a unique insight in specific use cases. Some examples for the latter include PhenoPath, which can include additional covariates in its model, Ouija, which returns a measure of uncertainty of each cell's position within the trajectory, and StemID, which can infer the directionality of edges within the trajectory. We feel that such methods should also be valued, and encourage their use provided that they are contained in a user friendly tool.

________________
# Online Methods
## Data and code availability
→ 
The processed real and synthetic datasets used in this study are deposited on Zenodo ([doi.org/10.5281/zenodo.1211533](https://doi.org/10.5281/zenodo.1211533[a])) [@cannoodt_goldstandardsinglecell_].

The main analysis repository is available at `r print_url("benchmark.dynverse.org")` and is divided into several experiments. Each experiment has its own set of scripts and results, each accompanied by an illustrated readme which can be easily browsed and explored on the github website.

The analysis scripts call several other R packages, of which an overview is available at `r print_url("dynverse.org")`. These packages include **dynwrap**, used to wrap the output of methods into the common trajectory model, **dyneval**, which contains the evaluation metrics, **dynguidelines**, the guidelines app, and **dynplot** for plotting trajectories.
←
## Trajectory inference methods
<!-- Overview -->
We gathered a list of →`r n_tools`← trajectory inference tools (`r ref("stable", "tools")`), by searching the literature for "trajectory inference" and "pseudotemporal ordering", and based on two existing lists found online: →`r print_url("https://github.com/seandavi/awesome-single-cell")`  [@davis_awesomesinglecelllistsoftware_2018] and  `r print_url("https://github.com/agitter/single-cell-pseudotime")` [@gitter_singlecellpseudotimeoverviewalgorithms_2018]←.
We welcome any contributions by creating an issue at →`r print_url("methods.dynverse.org")`←.

<!-- Inclusion criterion -->
Methods were excluded from the evaluation based on several criteria: `r glue::glue_collapse(glue::glue("({non_inclusion_reasons$footnote}) {non_inclusion_reasons$long}"), ", ")`. →The discussions on why these methods were excluded can be found at `r print_url("https://github.com/dynverse/dynmethods/issues?q=label:\"won't wrap\"")`. In the end, we included  `r n_tools_evaluated` tools in the evaluation, constituting `r n_methods_evaluated` methods.←

<!-- Components -->
The selected methods were dissected into a set of algorithmic components. Each component can have a significant impact on the performance, scalability, and output. Across all methods, these components can be broadly grouped into two stages; (i) conversion to a simplified representation using dimensionality reduction, clustering or graph building and (ii) ordering the cells along the simplified representation [@cannoodt_computational_2016]. Components are frequently shared between different algorithms. For example, minimal spanning trees (MST), used in the first single-cell RNA-seq trajectory inference methods [@trapnell_dynamics_2014], is shared by almost half of the methods we evaluated.
→
### Method wrappers
Each method was wrapped within docker and singularity containers (available at `r print_url("methods.dynverse.org")`). These containers are automatically built on both Singularity Hub (`r print_url("https://singularity-hub.org/")`) and Docker Hub (`r print_url("https://hub.docker.com/u/dynverse")`).
For each method, we wrote a wrapper script based on example scripts or tutorials provided by the authors (as mentioned in the respective wrapper scripts). This script reads in the input data (as described below), runs the method, and outputs the files required to construct a trajectory (as described below). We also created a script to generate an example dataset, which is used to automatically test every method using travis continuous integration (`r print_url("https://travis-ci.org/dynverse")`).

We used the github issues system to contact the authors of each method, and asked for feedback on the wrappers, the metadata, and the quality control. About one third of the authors responded, and we improved the wrappers based on their feedback. These discussions can be viewed on github: `r print_url('https://github.com/dynverse/dynmethods/issues?q=label:"method+discussion"')`.
←
### Method input
As input, we provided each method with either the raw count data (after cell and gene filtering) or normalised expression values, based on the description in the method documentation or from the study describing the method.
<!-- Prior information -->
A large portion of the methods requires some form of prior information (e.g. a start cell) in order to be executable. Other methods optionally allow to exploit certain prior information. Prior information can be supplied as a starting cell from which the trajectory will originate, a set of important marker genes, or even a grouping of cells into cell states. Providing prior information to a TI method can be both a blessing and a curse. In one way, prior information can help the method to find the correct trajectory among many, equally likely, alternatives. On the other hand, incorrect or noisy prior information can bias the trajectory towards current knowledge. Moreover, prior information is not always easily available, and its subjectivity can therefore lead to multiple equally plausible solutions, restricting the applicability of such TI methods to well studied systems.

The prior information was extracted from the →reference trajectory← as follows:

* **Start cells** The identity of one or more start cells. For both real and synthetic data, a cell was chosen which was the closest (in geodesic distance) to each milestone with only outgoing edges. For ties, one random cell was chosen. For cyclic datasets, a random cell was chosen.
* **End cells** The identity of one or more end cells. Similar as the start cells, but now for every state with only ingoing edges.
* **# end states** Number of terminal states. Number of milestones with only ingoing edges.
* **Grouping** For each cell a label to which state/cluster/branch it belongs. For real data, the states from the gold/silver standard. For synthetic data, each milestone was seen as one group, and cells were assigned to their closest milestone.
* **# branches** Number of branches/intermediate states. For real data, the number of states in the gold/silver standard. For synthetic data, the number of milestones.
→
* **Discrete time course** For each cell a time point from which it was sampled. If available, directly extracted from the reference trajectory, otherwise the geodesic distance from the root milestone was used. For synthetic data, four discrete timepoints were chosen, at which the cells were “sampled” to provide a time course information reflecting the one provided in real experiments.
* **Continuous time course** For each cell a time point from which it was sampled. For real data this was equal to the discrete time course, for synthetic data we used the internal simulation time of each simulator.
←
### Common trajectory model
Due to the absence of a common format for trajectory models, most methods return a unique set of output formats with few overlaps. →We therefore post-processed the output of each method into a common probabilistic trajectory model← (`r ref("sfig", "overview_wrapping", "a")`). This model consisted of three parts: (i) The milestone network represents the overall network topology, and contains edges between different milestones and the length of the edge between them. (ii) The milestone percentages contains, for each cell, its position between milestones, and sums for each cell to one. (iii) →The regions of delayed commitment, connections between three or more milestones. These must be explicitly defined in the trajectory model and per region, one milestone must be directly connected to all other milestones of the region.←

```{r}
add_sfig(
  raw_file("overview_wrapping_v2.svg", "03-methods"),
  "overview_wrapping", 
  "A common interface for TI methods.", 
  "a) The input and output of each TI method is standardised. As input, each TI method receives either raw or normalised counts, several parameters, and a selection of prior information. After its execution, a method uses one of the seven wrapper functions to transform its output to the common trajectory model. This common model then allows to perform common analysis functions on trajectory models produced by any TI method. b) The specific transformations performed by each of the wrapper functions is shown in more detail.", 
  13, 
  7.5
)
```

Depending on the output of a method, we used different strategies to convert the output to our model (`r ref("sfig", "overview_wrapping", "b")`). Special conversions are denoted by an \*, and will be explained in more detail below.

```{r}
methods_evaluated <- read_rds(result_file("methods_evaluated.rds", "03-methods"))
special_methods <- c("mfa", "dpt", "slice", "sincell", "calista", "urd")
# create functions to list the methods within a particular conversion category
conversion <- function(conversion_oi) {
  methods_evaluated %>% 
    filter(source %in% c("tool", "offtheshelf", "adaptation")) %>%
    filter(map_lgl(output, ~conversion_oi %in% .$outputs)) %>% 
    mutate(name = paste0(name, ifelse(id %in% special_methods, "\\*", ""))) %>% 
    pull(name) %>% 
    sort() %>%
    label_vector()
}
```

* **Type 1, direct:** `r conversion("trajectory")`. The wrapped method directly returned a network of milestones, the regions of delayed commitment, and for each cell is given to what extent it belongs to a milestone. →In some cases, this← indicates that additional transformations were required for the method not covered by any of the following output formats. →Some methods (`r conversion("branch_trajectory")`) returned a branch network instead of a milestone network, and this conversion was done by calculating the line graph of the branch network.←
* **Type 2, linear pseudotime:** `r conversion("linear_trajectory")`. The method returned a pseudotime, which is translated into a linear trajectory where the milestone network contains two milestones and cells are positioned between these two milestones.
* **Type 3, cyclical pseudotime:** `r conversion("cyclic_trajectory")`. The method returned a pseudotime, which is translated into a cyclical trajectory where the milestone network contains three milestones and cells are positioned between these three milestones. →These milestones were positioned at pseudotime 0, ⅓ and ⅔ .←
* **Type 4, end state probability:** `r conversion("end_state_probabilities")`. The method returned a pseudotime and for each cell and end state a probability for how likely a cell will end up in a certain end state. →This was translated into a star-shaped milestone network, with one starting milestone ($M_0$) and several outer milestones ($M_i$), with regions of delayed commitment between all milestones. The milestone percentage of a cell to one of the outer milestones was equal to $\textrm{pseudotime} \times \textrm{Pr}_{M_i}$ , the milestone percentage to the starting milestone was equal to $1-\textrm{pseudotime}$ .←
* **Type 5, cluster assignment:** `r conversion("cluster_graph")`. The method returned a milestone network, and an assignment of each cell to a specific milestone. Cells were positioned onto the milestones they are assigned to→, with milestone percentage equal to 1.←
* **Type 6, project onto nearest branch:** `r conversion("dimred_projection")`. The method returned a milestone network, and a dimensionality reduction of the cells and milestones. The cells were projected onto the closest nearest segment, thus determining the cells position along the milestone network. →If a method also returned a cluster assignment (type 5), we limited the projection of each cell to the closest edge connecting to the milestone of a cell. For these methods, we usually wrote two wrappers, one which included the projection, and one without.←
* **Type 7, cell graph:** `r conversion("cell_graph")`. The method returned a network of cells, 
and which cell-cell transitions are part of the 'backbone' structure. Backbone cells with degree $\neq$ 2 were regarded as milestones, and all other cells were placed on transitions between the milestones. →If methods did not return a distance between pairs of cells, the cells were uniformly positioned between the two milestones. Otherwise, we first calculated the distance between two milestones as the sum of the distances between the cells, and then divided the distance of each pair of cells with the total distance to get the milestone percentages.←

Special conversions were necessary for certain methods:

* **DPT:** We projected the cells onto the cluster network, consisting of a central milestone (this cluster contains the cells which were assigned to the "unknown" branch) and three terminal milestones, each corresponding to a tip point. →This was then processed as a type 1, direct,  method.←
* **Sincell:** To constrain the number of milestones this method creates, we merged two cell clusters iteratively until the percentage of leaf nodes was below a certain cutoff, default at 25%.  →This was then processed as a type 7, cell graph, method.←
* **SLICE:** As discussed in the vignette of SLICE, we ran principal curves one by one for every edge detected by SLICE. →This was then processed as a type 1, direct, method.←
* **mfa:** We used the branch assignment as state probabilities, which together with the global pseudotime were processed →as a type 4, end state probabilities, method.←
→
* **CALISTA:** We assigned the cells to the branch at which the sum of the cluster probabilities of two connected milestones was the highest. The cluster probabilities of the two selected milestones were then used as milestone percentages. This was then processed as a type 1, direct, method.
* **URD:** We extracted the pseudotime of a cell within each branch using the y positions in the tree layout. This was then further processed as a type 1, direct, method.

More information on how each method was wrapped can be found within the comments of each wrapper script, available at `r print_url("https://github.com/dynverse/dynmethods")`.←

→
### Control methods
We added several control positive and negative control methods: **random**, which generates a random network with 50 milestones according to a Barabási–Albert model, **identity**, which returns the reference trajectory, and **shuffle**, which will place the cells randomly on the reference milestone network.
### Off-the-shelf methods
For baseline performance, we added several "off-the-shelf" TI methods which can be run using just a couple of commands in R.←

* **Component 1:** This method returns the first component of a PCA dimensionality reduction as a linear trajectory. Although conceptually very simple, it has been used in certain studies [@kouno_temporaldynamicstranscriptional_2013; @zeng_pseudotemporalorderingsingle_2017].
* **Angle:** Similar to the previous method, this method computes the angle with respect to the origin in a two-dimensional PCA, and uses this angle as pseudotime for a cyclical trajectory.
→
* **Periodic PrinCurve:** Periodic principal curves, using the R princurve package
* **MST:** Dimensionality reduction with PCA, followed by clustering using the R mclust package, after which the clusters are connected using a minimal spanning tree. The cells are then projected onto the edges of the MST.
←

## Real datasets
We gathered real datasets by searching for "single-cell" at the Gene Expression Omnibus and selecting those datasets in which the cells are sampled from different stages in a dynamic process (`r ref("stable", "datasets")`). The scripts to download and process these datasets are available on our repository (`r print_url("https://github.com/dynverse/dynbenchmark/tree/master/scripts/01-datasets")`). Whenever possible, we preferred to start from the raw counts data. These raw counts were all normalised and filtered using a common pipeline, discussed later. →Some original datasets contained more than one trajectory, in which case we split the dataset into its separate connected trajectory, but also generated several combinations of connected trajectories to include some datasets with disconnected trajectories in the evaluation. In the end, we included `r sum(list_datasets()$source == "real")` for this evaluation.←

→For each dataset, we extracted a reference trajectory, consisting of two parts: the cellular grouping (milestones) and the connections between these groups (milestone network). The cellular grouping was provided by the authors of the original study, and we classified it as a gold standard when it was created independently from the expression matrix (such as from cell sorting, the origin of the sample or the time it was sampled) or silver standard otherwise (usually by clustering the expression values). To connect these cell groups, we used the original study to determine the network which the authors validated or otherwise found to be the most likely. In the end, each group of cells was placed on a milestone, having percentage = 1 for that particular milestone. The known connections between these groups were used to construct the milestone network. If there was time data available, we used this as the length of the edge, otherwise we set all the lengths equal to one. ←

```{r}
add_stable(
  result_file("metadata.xlsx","01-datasets/01-real"), 
  "datasets", 
  "Real datasets used in this study."
)
```
## Synthetic datasets
→
To generate synthetic datasets, we used four different synthetic data simulators:

* **dyngen:** Simulations of gene regulatory networks, available at `r print_url("https://github.com/dynverse/dyngen")`
* **dyntoy:** Random gradients of expression in the reduced space, available at `r print_url("https://github.com/dynverse/dyntoy")`
* **PROSSTT:** Expression is sampled from a linear model which depends on pseudotime [@papadopoulosPROSSTTProbabilisticSimulation2018]
* **Splatter:** Simulations of non-linear paths between different expression states [@zappia_splattersimulationsinglecell_2017]

For every simulator, we took great care to make the datasets as realistic as possible. To do this, we extracted several parameters from all real datasets. We calculated the number of differentially expressed features within a trajectory using a wilcoxon test between every pair of cell groups. These values were corrected for multiple testing using the Benjamini-Hochberg procedure (FDR < 0.05) and we required that a gene was expressed in at least 5% of cells, and had at least a fold-change of 2. We also calculated several other parameters, such as drop-out rates and library sizes using the Splatter package [@zappia_splattersimulationsinglecell_2017]. These parameters were then given to the simulators when applicable, as described for each simulator below. To select a diverse set of $k$ reference real datasets, we clustered all the parameters of each real dataset, and used the reference real datasets at the cluster centers from a pam clustering (implemented in the R cluster package) to generate synthetic data.
←
### dyngen

```{r}
design_dyngen <- read_rds(result_file("design_dyngen.rds", "01-datasets/02-synthetic"))
```

The dyngen ( `r print_url("https://github.com/dynverse/dyngen")`) workflow to generate synthetic data is based on the well established workflow used in the evaluation of network inference methods [@schaffter_genenetweaversilicobenchmark_2011; @marbach_wisdom_2012] and consists of four main steps: network generation, simulation, gold standard extraction and simulation of the scRNA-seq experiment. At every step, we tried to mirror real regulatory networks, while keeping the model simple and easily extendable. →We simulated a total of `r nrow(design_dyngen)` datasets, with `r length(unique(design_dyngen$modulenet_name))` different topologies.←
#### Network generation

One of the main processes involved in cellular dynamic processes is gene regulation, where regulatory cascades and feedback loops lead to progressive changes in expression and decision making. The exact way a cell choses a certain path during its differentiation is still an active research field, although certain models have already emerged and been tested in vivo. One driver of bifurcation seems to be mutual antagonism, where genes [@xu_regulationbifurcatingcell_2015] strongly repress each other, forcing one of the two to become inactive [@graf_forcingcellschange_2009]. Such mutual antagonism can be modelled and simulated [@wang_quantifyingwaddingtonlandscape_2011; @ferrell_bistabilitybifurcationswaddington_2012]. Although such a two-gene model is simple and elegant, the reality is frequently more complex, with multiple genes (grouped into modules) repressing each other [@yosef_dynamicregulatorynetwork_2013].

→To simulate certain trajectory topologies, we therefore designed module networks in which the cells follow a particular trajectory topology given certain parameters. Two module networks generated linear trajectories (linear and linear long), one generated a bifurcation, one generated a convergence, one generated a multifurcation (trifurcating), two generated a tree (consecutive bifurcating and binary tree), one generated an acyclic graph (bifurcating converging), (bifurcating and converging), one generated a complex fork (trifurcating), one generated a rooted tree (consecutive bifurcating) and two generated simple graph structures (bifurcating loop and bifurcating cycle). The structure of these module networks is available at `r print_url("https://github.com/dynverse/dyngen/tree/master/inst/ext_data/modulenetworks")`.←

From these module networks we generated gene regulatory networks in two steps: the main regulatory network was first generated, and extra target genes from real regulatory networks  were added. For each dataset, we used the same number of genes as were differentially expressed in the real datasets. 5% of the genes were assigned to be part of the main regulatory network, and were randomly distributed among all modules (with at least one gene per module). We sampled edges between these individual genes (according to the module network) using a uniform distribution between 1 and the number of possible targets in each module. To add additional target genes to the network, we assigned every regulator from the network to a real regulator in a real network (from regulatory circuits [@marbach_tissuespecificregulatorycircuits_2016]), and extracted for every regulator a local network around it using personalized pagerank (with damping factor set to 0.1), as implemented in the `page_rank` function of the *igraph* package.

#### Simulation of gene regulatory systems using thermodynamic models

To simulate the gene regulatory network, we used a system of differential equations similar to those used in evaluations of gene regulatory network inference methods [@marbach_wisdom_2012]. In this model, the changes in gene expression ($x_i$) and protein expression ($y_i$) are modeled using ordinary differential equations [@schaffter_genenetweaversilicobenchmark_2011] (ODEs):

$$
\begin{aligned}
\label{eq:mrna_ode}
\frac{dx_i}{dt} &= \underbrace{m \times f(y_1, y_2, ...)}_\text{production} - \underbrace{\lambda \times x_i}_\text{degradation}
\end{aligned}
$$
$$
\begin{aligned}
\label{eq:prot_ode}
\frac{dy_i}{dt} &= \underbrace{r \times x_i}_\text{production} - \underbrace{\Lambda \times y_i}_\text{degradation}
\end{aligned}
$$

where $m$, $\lambda$, $r$ and $\Lambda$ represent production and degradation rates, the ratio of which determines the maximal gene and protein expression. The two types of equations are coupled because the production of protein $y_i$ depends on the amount of gene expression $x_i$, which in turn depends on the amount of other proteins through the activation function $f(y_1, y_2, ...)$.

The activation function is inspired by a thermodynamic model of gene regulation, in which the promoter of a gene can be bound or unbound by a set of transcription factors, each representing a certain state of the promoter. Each state is linked with a relative activation $\alpha_j$, a number between 0 and 1 representing the activity of the promoter at this particular state. The production rate of the gene is calculated by combining the probabilities of the promoter being in each state with the relative activation:

$$
\begin{aligned}
\label{eq:input_function_probabilities}
f(y_1, y_2, ..., y_n) = \sum_{j \in \{0, 1, ..., n^2\}} \alpha_j \times P_j
\end{aligned}
$$

The probability of being in a state is based on the thermodynamics of transcription factor binding. When only one transcription factor is bound in a state:
$$
\begin{aligned}
P_j \propto \nu = \left(\frac{y}{k}\right)^{n}
\end{aligned}
$$

where the hill coefficient $n$ represents the cooperativity of binding and $k$ the transcription factor concentration at half-maximal binding. When multiple regulators are bound:
$$
\begin{aligned}
P_j \propto \nu =  \rho \times \prod_j \left(\frac{y_j}{k_j}\right)^{n_j}
\end{aligned}
$$

where $\rho$ represents the cooperativity of binding between the different transcription factors.

$P_i$ is only proportional to $\nu$ because $\nu$ is normalized such that $\sum_{i} P_i = 1$.

To each differential equation, we added an additional stochastic term: 
$$
\begin{aligned}
\frac{dx_i}{dt} = m \times f(y_1, y_2, ...) - \lambda \times x_i &+ \eta \times \sqrt{x_i} \times \Delta W_t \\
\frac{dy_i}{dt} = r \times x_i - \Lambda \times y_i &+ \eta \times \sqrt{y_i} \times \Delta W_t
\end{aligned}
$$

with $\Delta W_t \sim \mathcal{N}(0, h)$.

Similar to [@schaffter_genenetweaversilicobenchmark_2011], we sample the different parameters from random distributions, given in `r ref("stable", "samplers")`.

```{r}
add_stable(
  read_rds(result_file("samplers.rds","01-datasets/02-synthetic")), 
  "samplers", 
  "Distributions from which each parameter in the thermodynamic model was sampled."
)
```

We converted each ODE to an SDE by adding a chemical Langevin equation, as described in [@schaffter_genenetweaversilicobenchmark_2011]. These SDEs were simulated using the Euler–Maruyama approximation, with time-step $h = 0.01$ and noise strength $\eta = 8$. The total simulation time varied between 5 for linear and bifurcating datasets, 10 for consecutive bifurcating, trifurcating and converging datasets, 15 for bifurcating converging datasets and 30 for linear long, cycle and bifurcating loop datasets. The burn-in period was for each simulation 2. Each network was simulated 32 times.

#### Simulation of the single-cell RNA-seq experiment

For each dataset we sampled the same number of cells as were present in the reference real dataset, limited to the simulation steps after burn-in. →These cells were sampled uniformly across the different steps of the 32 simulations. ← Next, we used the Splatter package [@zappia_splattersimulationsinglecell_2017] to estimate the different characteristics of a real dataset, such as the distributions of average gene expression, library sizes and dropout probabilities. We used Splatter to simulate the expression levels $\lambda_{i,j}$ of housekeeping genes $i$ (to match the number of genes in the reference dataset) in every cell $j$. These were combined with the expression levels of the genes simulated within a trajectory. Next, true counts were simulated using $Y'_{i,j} \sim \text{Poisson}(\lambda_{i,j})$. Finally, we simulated dropouts by setting true counts to zero by sampling from a Bernoulli distribution using a dropout probability $\pi^D_{i,j} =\frac{1}{1+e^{-k(\text{ln}(\lambda_{i,j})-x_0)}}$. →Both $x_0$ (the midpoint for the dropout logistic function) and $k$ (the shape of the dropout logistic function) were estimated by Splatter.←

This count matrix was then filtered and normalised using the pipeline described below.

#### Gold standard extraction

Because each cellular simulation follows the trajectory at its own speed, knowing the exact position of a cell within the trajectory topology is not straightforward. Furthermore, the speed at  which simulated cells make a decision between two or more alternative paths is highly variable. →We therefore first constructed a backbone expression profile for each branch within the trajectory. To do this, we first defined in which order the expression of the modules is expected to change, and then generated a backbone expression profile in which the expression of these modules increases and decreases smoothly between 0 and 1.  We also smoothed the expression in each simulation using a rolling mean with a window of 50 time steps, and then calculated the average module expression along the simulation. ← We used dynamic time warping, implemented in the dtw R package [@giorgino_computingvisualizingdynamic_2009; @tormene_matchingincompletetime_2009], with an open end to align a simulation to all possible module progressions, and then picked the alignment which minimised the normalised distance between the simulation and the backbone. In case of cyclical trajectory topologies, the number of possible milestones a backbone could progress through was limited to 20.

→
### dyntoy
```{r}
design_dyntoy <- read_rds(result_file("design_dyntoy.rds", "01-datasets/02-synthetic"))
```

For more simplistic data generation ("toy" datasets), we created the dyntoy workflow (`r print_url("https://github.com/dynverse/dyntoy")`) . We created `r length(unique(design_dyntoy$topology_model))` topology generators (described below), and with   `r nrow(design_dyntoy)/length(unique(design_dyntoy$topology_model))` datasets per generator, this lead to a total of `r nrow(design_dyntoy)` datasets.

We created a set of topology generators:

* Linear and cyclic, with number of milestones $\sim B(10, 0.25)$
* Bifurcating and converging, with four milestones
* Binary tree, with number of branching points $\sim U(3, 6)$
* Tree, with number of branching points $\sim U(3, 6)$ and maximal degree $\sim U(3, 6)$

For more complex topologies we first calculated a random number of "modifications" $\sim U(3, 6)$ and a $\textit{deg}_{\textit{max}} \sim B(10, 0.25) + 1$. For each type of topology, we defined what kind of modifications are possible: divergences, loops, convergences and divergence-convergence. We then iteratively constructed the topology by uniformly sampling from the set of possible modifications, and adding this modification to the existing topology. For a divergence, we connected an existing milestone to a number of a new milestones. Conversely, for a convergence we connected a number of new nodes to an existing node. For a loop, we connected two existing milestones with a number of milestones in between. Finally for a divergence-convergence we connected an existing milestone to several new milestones which again converged on a new milestone. The number of nodes was sampled from $\sim B(\textit{deg}_{\textit{max}} - 3, 0.25) + 2$

* Looping, allowed loop modifications
* Diverging-converging, allowed divergence and converging modifications
* Diverging with loops, allowed divergence and loop modifications
* Multiple looping, allowed looping modifications
* Connected, allowed looping, divergence and convergence modifications
* Disconnected, number of components sampled from $\sim B(5, 0.25) + 2$, for each component we randomly chose a topology from the ones listed above

After generating the topology, we sampled the length of each edge $\sim U(0.5, 1)$. We added regions of delayed commitment to a divergence by in half of the cases. We then placed the number of cells (same number as from the reference real dataset), on this topology uniformly, based on the length of the edges in the milestone network.

For each gene (same number as from the reference real dataset), we calculated the Kamada-Kawai layout in 2 dimensions, with edge weight equal to the length of the edge. For this gene, we then extracted for each cell a density value using a bivariate normal distribution with $\mu \sim U(x_{\textit{min}}, x_{\textit{min}})$ and $\sigma \sim U(x_{\textit{min}}/10, x_{\textit{min}}/8)$. We used this density as input for a zero-inflated negative binomial distribution with $\mu ~ U(100, 1000) \times \textit{density}$, $k ~ U(\mu / 10, \mu / 4)$ and $pi$ from the parameters of the reference real dataset, to get the final count values.

This count matrix was then filtered and normalised using the pipeline described below.
### PROSSTT
```{r}
design_prosstt <- read_rds(result_file("design_dyntoy.rds", "01-datasets/02-synthetic"))
```

PROSSTT is a recent data simulator [@papadopoulosPROSSTTProbabilisticSimulation2018], which simulates expression using linear mixtures of expression programs and random walks through the trajectory. We used 5 topology generators from dyntoy (linear, bifurcating, multifurcating, binary tree and tree), and simulated for each topology generator 10 datasets using different reference real datasets. However, due to frequent crashes of the tool, only 19 datasets created output and were thus used in the evaluation.

Using the `simulate_lineage` function, we simulated the lineage expression, with parameters $\a \sim U(0.01, 0.1)$, $\textit{branch-tol}_{\textit{intra}} \sim U(0, 0.9)$ and $\textit{branch-tol}_{\textit{inter}} \sim U(0, 0.9)$. These parameter distributions were chosen very broad so as to make sure both easy and difficult datasets are simulated. After simulating base gene expression with `simulate_base_gene_exp`, we used the `sample_density` function to finally simulate expression values of a number of cells (the same as from the reference real dataset), with $\alpha \sim \textit{Lognormal}$ ($\mu = 0.3$ and $\sigma = 1.5$) and $\beta \sim \textit{Lognormal}$ ($\mu = 2$ and $\sigma = 1.5$). Each of these parameters were centered around the default values of PROSSTT, but with enough variability to ensure a varied set of datasets.

This count matrix was then filtered and normalised using the pipeline described below.
### Splatter
Splatter [@zappia_splattersimulationsinglecell_2017] simulates expression values by constructing non-linear paths between different states, each having a distinct expression profile. We used 5 topology generators from dyntoy (linear, bifurcating, multifurcating, binary tree and tree), and simulated for each topology generator 10 datasets using different reference real datasets, leading to a total of 50 datasets.

We used the `splatSimulatePaths` function from Splatter to simulate datasets, with number of cells and genes equal to those in the reference real dataset, and with parameters  $\textit{nonlinearProb}$, $\textit{sigmaFac}$ and $\textit{skew}$ all sampled from $U(0, 1)$.

This count matrix was then filtered and normalised using the pipeline described below.

←
## Dataset filtering and normalisation
We used a standard single-cell RNA-seq preprocessing pipeline which applies parts of the scran and scater Bioconductor packages [@lun_stepbystepworkflowlowlevel_2016]. The advantages of this pipeline is that it works both with and without spike-ins, and includes a harsh cell filtering which looks at abnormalities in library sizes, mitochondrial gene expression, and number of genes expressed using median absolute deviations (→which we set to 3←). We required that a gene was expressed in at least 5% of the cells, and that it should have an average expression higher than 0.02. Furthermore, we used the pipeline to select the most highly variable genes, using a false discovery rate of 5% and a biological component higher than 0.5. As a final filter, we removed both all-zero genes and cells until convergence.
## Evaluation metrics
The importance of using multiple metrics to compare complex models has been stated repeatedly [@norel_selfassessmenttrap_2011]. →Furthermore, a trajectory is a model with multiple layers of complexity, which calls for several metrics each assessing a different layer. We therefore defined several possible metrics for comparing trajectories, each investigating different layers. These are all discussed in `r ref("snote", "metrics")` along with examples and robustness analyses when appropriate.

Next, we created a set of rules to which we think a good trajectory metric should conform, and tested this empirically for each metric by comparing scores before and after perturbing a dataset (`r ref("snote", "metrics")`). Based on this analysis, we chose 4 metrics for the evaluation, each assessing a different aspect of the trajectory: (i) the `r label_metric("him")` measures the topological similarity, (ii) the   `r label_metric("F1_branches")`compares the branch assignment, (iii) the  `r label_metric("correlation")` assesses the similarity in pairwise cell-cell distances and thus the cellular positions, and (iv) the  `r label_metric("featureimp_wcor")` looks at whether similar important features (genes) are found in both the reference dataset and the prediction.

### The Hamming-Ipsen-Mikhailov metric
The Hamming-Ipsen-Mikhailov (HIM) [@jurman_HIM_Glocal_Metric_2015] uses the two weighted adjacency matrices of the milestone networks as input (weighted by edge length). It is a linear combination of the normalised Hamming distance, which gives an indication of the differences in edge lengths, and the normalised Ipsen-Mikhailov distance, which assesses the similarity in degree distributions. The latter has a parameter $\gamma$, which was fixed at 0.1 as to make the scores comparable between datasets. We illustrate the metric and discuss alternatives in `r ref("snote", "metrics")`.

### The F1 between branch assignments
To compare branch assignment, we borrowed the $F_1$ metric from the field of expression biclustering [@saelens_comprehensiveevaluationmodule_2018]. To calculate this metric, we first calculated the similarity of all pairs of branches between the two trajectories using the Jaccard similarity. Next, we define the $\textit{Recovery}$ (respectively $\textit{Relevance}$) as the average maximal similarity of all branches in the reference dataset (respectively $\textit{prediction}$). The `r label_metric("F1_branches")` was then defined as the harmonic mean between $\textit{Recovery}$ and $\textit{Relevance}$. We illustrate this metric further in  `r ref("snote", "metrics")`.
←
### Correlation between geodesic distances
→When the position of a cell is the same in both the reference and the prediction, its relative distances to all other cells in the trajectory should also be the same. This observation is the basis for the `r label_metric("correlation")` metric. 
To calculate the `r label_metric("correlation")`, we first sampled 100 waypoint cells in both the prediction and the reference dataset, using stratified sampling between the different milestones, edges and regions of delayed commitment, weighted by the number of cells in each collection. We then calculated the geodesic distances between the union of waypoint cells from both datasets and all other cells. The calculation of the geodesic distance depended on the location of the two cells within the trajectory, further discussed in `r ref("snote", "metrics")` and was weighted by the length of the edge in the milestone network. Finally, the `r label_metric("correlation")` was defined as the Spearman rank correlation between the distances of both datasets. We illustrate the metric and assess the effect of the number of waypoint cells in `r ref("snote", "metrics")`.
### The correlation between important features
The `r label_metric("featureimp_wcor")` assesses whether similar dynamically differentially expressed features are found between the prediction and model. To calculate this metric, we used Random Forest regression (implemented in the R ranger package), to predict expression values of each gene, based on the geodesic distances of a cell to each milestone. We then extract feature importance values for each feature, and calculated the similarity of the feature importances using a weighted Pearson correlation, weighted by the feature importance in the reference dataset as to give more weight to differentially expressed features then non-differentially expressed features. As hyperparameters we set the number of trees to 10000, and the number of features on which can be split to 1% of all available features. We illustrate this metric and assess the effect of its hyperparameters in `r ref("snote", "metrics")`.

### Aggregation of scores
To rank methods, we needed to aggregate the different scores on two levels: across datasets and across different metrics. This aggregation strategy is explained into more detail in `r ref("snote", "metrics")`.

To make sure easy and difficult datasets have equal influence on the final score, we first normalised the scores on each dataset across the different methods. We shifted and scaled the scores to $\sigma = 1$ and $\mu = 0$, and then applied the unit probability density function on these values to get the scores back into the $[0, 1]$ range.

Next, we calculated for each metric, trajectory type and dataset source an average score using an arithmetic mean.

Next, we aggregated the scores across trajectory types and dataset source (real/gold, real/silver, synthetic/dyngen, synthetic/splatter, synthetic/prosstt and synthetic/dyntoy), so that each of these had an equal weight on the final score:

1. Calculate a score for each method, dataset source and trajectory type
←
## Benchmark
### Method execution
Each execution of a method on a dataset was performed in a separate task as part of a gridengine job. Each task was allocated one CPU core of an Intel(R) Xeon(R) CPU E5-2665 @ 2.40GHz, and one R session was started for each task. During the execution of a method on a dataset, if the time limit (>6h) or memory limit (32GB) was exceeded, or an error was produced, a zero score was returned for that execution. If a method consistently generated errors on this dataset across all replicates, the error is called “data-specific”, otherwise “stochastic”. The environment variable `R_MAX_NUM_DLLS` set to 500. The `base::set.seed` was overridden in order to prevent stochastic TI methods from pretending to be deterministic and/or robust. Timings of methods were measured for different steps along the executions, including preprocessing, postprocessing, each of the different metrics, and the method itself.
### Effect of prior information
To assess the effect of prior information on the performance of a method, we compared the performance of methods which do or do not require the prior information using a two-tailed Mann–Whitney U test. P-values were controlled for multiple testing using Benjamini-Hochberg correction.
### Effect of algorithm components
To assess the effect of algorithm components on the performance of a method, we used (1) a two-tailed Mann–Whitney U test, as implemented in the `wilcox.test` R function, and (2) increase in node purity importance scores using random forest classification, as implemented in the R randomForest package, predicting whether a method scored better than the median score. To evaluate whether a method increased or decreased performance, we used the estimate of the location parameter of the Mann–Whitney U test. We only investigated algorithm components which were part of at least 4 different methods in our evaluation study. P-values were controlled for multiple testing using Benjamini-Hochberg correction.
→
## Scalability[b]
To assess the scalability of each method, we started from 5 real datasets, selected using the centers from a k-medoids as discussed in the synthetic data generation section. We up or down-scaled these datasets … OKAY I GIVE UP generate_dataset.R

## [c]Stability
←
## Method quality control

We created a transparent scoring scheme to check the quality of each method based on several existing tool quality and programming guidelines in literature and online (`r ref("stable", "qc_scoresheet")`). The goal of this quality control is in the first place is to stimulate the improvement of current methods, and the development of user and developer friendly new methods. The quality control assessed 6 categories, each looking at several aspects, which are further divided into individual items. The availability category checks whether the method is easily available, whether the code and dependencies can be easily installed, and how the method can be used. The code quality assesses the quality of the code both from a user perspective (function naming, dummy proofing and availability of plotting functions) and a developer perspective (consistent style and code duplication). The code assurance category is frequently overlooked, and checks for code testing, continuous integration [@beaulieu-jones_reproducibility_2017] and an active support system. The documentation category checks the quality of the documentation, both externally (tutorials and function documentation) and internally (inline documentation). The behaviour category assesses the ease by which the method can be run, by looking for unexpected output files and messages, prior information and how easy the trajectory model can be extracted from the output. Finally, we also assessed certain aspects of the study in which the method was proposed, such as publication in a peer-reviewed journal, the number of dataset in which the usefulness of the method was shown, and the scope of method evaluation in the paper.

Each aspect was further assigned to one or more applications, based on whether it influenced the user friendliness of the tool (such as code availability, good documentation or contains plotting functions) and/or the developer friendliness of the tool (such as unit testing, inline documentation and a clear licensing).

Each quality aspect received a weight depending on how frequently it was found in several papers and online sources which discuss tool quality (`r ref("stable", "qc_scoresheet")`). This was to make sure that more important aspects, such as the open source availability of the method, outweighed other aspects, such as the availability of a graphical user interface. Within each aspect, we also assigned a weight to the individual questions being investigated (`r ref("stable", "qc_scoresheet")`). For calculating the final score, we weighed each of the six categories equally.
## Trajectory types
We classified all possible trajectory topologies into distinct trajectory types, based on topological criteria (`r ref("fig", "ti_evaluation_overview", "c")`). These trajectory types start from the most general trajectory type, a disconnected directed graph, and move down (within a directed acyclic graph structure), progressively becoming more simple until the two basic types: linear and cyclical. A disconnected directed graph is a graph in which only one edge can exist between two nodes. A directed graph is a disconnected graph in which all nodes are (indirectly) connected. A directed acyclic graph is a graph containing no cycles. A rooted tree is a directed acyclic graph containing no convergences (no nodes with in-degree higher than 1). A rooted binary tree is a rooted tree with every node having maximally two outgoing edges. A convergence is a directed acyclic graph in which only one node has a degree larger than one and this same node has an indegree of one. A multifurcation is a rooted tree in which only one node has a degree larger than one. A bifurcation is a multifurcation in which only one node has a degree equal to 3. A directed linear graph is a graph in which no node has a degree larger than 3. Finally, a directed cycle is a directed graph in which no node has a degree larger than 3 but contains one cycle. For every directed trajectory type, a corresponding undirected trajectory type can also be defined, with directed acyclic graphs and directed graphs merging to undirected graph, and convergence and bifurcation merging to a simple fork. In most cases, a method which was able to detect a complex trajectory type, was also able to detect less complex trajectory types, with →the exceptions being being DPT (limited to bifurcating trajectories), and STEMNET and FateID (unable to return linear trajectories).←

# Acknowledgements
We would like to thank all the original authors of the methods for their feedback and improvements on the method wrappers.

<!--
Tips for working in this document:
Go to tools -> preferences and remove Use smart quotes
Enable automatic zotero syncing to bibtex
Make sure to pin the zotero bibtex references after first using them!
-->

[a]Update the doi when reuploaded
[b]I will try to do my best in adding this if you don't do it first Robrecht
[c]I will try to do my best in adding this if you don't do it first Robrecht