### 2018-11-12 manuscript/supplementary.Rmd
update paper & supplementary

---
title: "Supplementary material for: \nA comparison of single-cell trajectory inference methods"
output: 
  dynbenchmark::pdf_manuscript:
      render_change: FALSE
---

```{r setup, include=FALSE}
library(dynbenchmark)
library(tidyverse)
```

```{r}
supplementary <- read_rds("supplementary.rds")
sfigs <- supplementary$sfigs
stables <- supplementary$stables
refs <- supplementary$refs
```

# Supplementary figures
```{r, results="asis"}
sfigs %>% 
  left_join(refs %>% select(ref_id, name), "ref_id") %>% 
  arrange(name) %>% 
  select(-name) %>% 
  pmap(dynbenchmark:::plot_fig) %>% modify(paste0, "\n\n") %>% walk(cat)
```

\clearpage

# Supplementary tables

```{r, results="asis"}
stables %>% 
  left_join(refs %>% select(ref_id, name), "ref_id") %>% 
  arrange(name) %>% 
  select(-name) %>% 
  pmap(dynbenchmark:::show_table) %>% modify(paste0, "\n\n") %>% walk(cat)
```

\clearpage

# Supplementary note 1

```{r}
# dynbenchmark::knit_nest("../results/02-metrics/")
```