### 2018-11-14 README.Rmd

---
output: dynbenchmark::github_markdown_nested
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = FALSE, echo = FALSE)
```

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(dynbenchmark)
```

[![Build Status](https://api.travis-ci.org/dynverse/dynbenchmark.svg)](https://travis-ci.org/dynverse/dynbenchmark)
![Lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg) <a href = "package/man/figures/logo.svg"><img src="package/man/figures/logo.png" align="right" /></a> [![doi](https://zenodo.org/badge/doi/10.1101/276907.svg)](https://doi.org/10.1101/276907)

# Benchmarking trajectory inference methods

This repo contains the scripts to reproduce the manuscript

```{r}
get_altmetric_badge <- function() {
  curl::curl_fetch_memory("https://api.altmetric.com/v1/doi/10.1101/276907")$content %>%
    rawToChar() %>% 
    jsonlite::fromJSON() %>% 
    purrr::pluck("images") %>%
    purrr::pluck("medium") %>% 
    paste0("&style=bar")
}
```

> A comparison of single-cell trajectory inference methods: towards more accurate and robust tools  
<strong> Wouter Saelens\* </strong> <a href='https://orcid.org/0000-0002-7114-6248'><img src='https://github.com/dynverse/dynmethods/raw/master/man/figures/orcid_logo.svg?sanitize=true' height='16'></a> <a href='https://github.com/zouter'><img src='https://github.com/dynverse/dynmethods/raw/master/man/figures/github_logo.png' height='16'></a>,
<strong> Robrecht Cannoodt\* </strong> <a href='https://orcid.org/0000-0003-3641-729X'><img src='https://github.com/dynverse/dynmethods/raw/master/man/figures/orcid_logo.svg?sanitize=true' height='16'></a> <a href='https://github.com/rcannood'><img src='https://github.com/dynverse/dynmethods/raw/master/man/figures/github_logo.png' height='16'></a>,
Helena Todorov <a href='https://github.com/Helena-todd'><img src='https://github.com/dynverse/dynmethods/raw/master/man/figures/github_logo.png' height='16'></a>,
<em> Yvan Saeys </em> <a href='https://github.com/saeyslab'><img src='https://github.com/dynverse/dynmethods/raw/master/man/figures/github_logo.png' height='16'></a>  
[bioRxiv:276907](https://www.biorxiv.org/content/early/2018/03/05/276907) [doi:10.1101/276907](https://doi.org/10.1101/276907) [![altmetric](`r get_altmetric_badge()`)](https://altmetric.com/details/33972849)

(Note that this preprint is currently very outdated, see below for more recent results)

## Experiments

From start to finish, the repository is divided into several experiments, each with their own scripts and results. These are accompanied by documentation using github readmes and can thus be easily explored by going to the appropriate folders:

```{r results = 'asis'}
extract_scripts_documentation("scripts", recursive = FALSE) %>% 
  mutate(
    scripts = location,
    results = map_chr(scripts, dynbenchmark::link_to_results)
  ) %>% 
  mutate(
    ix = case_when(is.na(ix) ~ "", TRUE ~ as.character(ix)),
    id = label_long(id),
    scripts = paste0("[\U1F4C4\U27A1](", scripts, ")"),
    results = case_when(is.na(results) ~ "", TRUE ~ paste0("[\U1F4CA\U27A1](", results, ")"))
  ) %>% 
  select(`\\#` = ix, id, scripts, results) %>% 
  knitr::kable()
```

We also have several additional subfolders:

* [Manuscript](manuscript): Source files for producing the manuscript.
* [Package](package): An R package with several helper functions.
* [Raw](raw): Files generated by hand, such as figures and spreadsheets.
* [Derived](derived): Intermediate data files produced by the scripts. These files or not git committed.

## Guidelines

Based on the results of the benchmark, we provide context-dependent user guidelines, [available as a shiny app](https://github.com/dynverse/dynguidelines). This app is integrated within the [dyno pipeline](https://github.com/dynverse/dyno), which also includes the wrappers used in the benchmarking and other packages for visualising and interpreting the results.

[![dynguidelines](https://github.com/dynverse/dynguidelines/raw/master/man/figures/demo.gif)](https://github.com/dynverse/dynguidelines)

## Datasets

The benchmarking pipeline generates (and uses) the following datasets:

* **Gold standard single-cell datasets**, both real and synthetic, used to evaluated the trajectory inference methods [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.1211533.svg)](https://doi.org/10.5281/zenodo.1211533)

```{r, echo = FALSE, message=FALSE}
set.seed(3)

datasets <- list_datasets() %>% group_by(source) %>% sample_n(1) %>% pull(id) %>% load_datasets()

datasets$plot_dimred <- mapdf(datasets, ~plot_dimred(., dimred = dyndimred::dimred_landmark_mds) + ggtitle(.$source) + theme(plot.title = element_text(size = 10)))
p <- patchwork::wrap_plots(datasets$plot_dimred, ncol = nrow(datasets))
ggplot2::ggsave("package/man/figures/datasets.png", width = nrow(datasets)*3, height = 1*3)
```

![datasets](package/man/figures/datasets.png)

* **The performance of methods** used for the [results overview figure](`r dynbenchmark::link_to_results("scripts/08-summary/overview.pdf")`) and the [dynguidelines](https://www.github.com/dynverse/dynguidelines) app. _Not yet available_, but can be accessed in the results folder.

* **General information about trajectory inference methods**, available as a data frame in `dynmethods::methods`

## Methods

All wrapped methods are wrapped as both docker and singularity containers. These can be easily run using [*dyn*methods](https://github.com/dynverse/dynmethods).

## Installation

dynbenchmark has been tested using R version 3.5.1 on Linux. While running the methods also works on on Windows and Mac (see [dyno](https://github.com/dynverse/dyno)), running the benchmark is currently not supported on these operating system, given that a lot of commands are linux specific.

In R, you can install the dependencies of dynbenchmark from github using:

``` r
# install.packages("devtools")
devtools::install_github("dynverse/dynbenchmark")
```

This will install several other "dynverse" packages. Depending on the number of R packages already installed, this installation should take approximately 5 to 30 minutes.

On Linux, you will need to install udunits and ImageMagick:

* Debian / Ubuntu / Linux Mint: `sudo apt-get install libudunits2-dev imagemagick`
* Fedora / CentOS / RHEL: `sudo dnf install udunits2-devel ImageMagick-c++-devel`

[Docker](https://docs.docker.com/install) has to be installed to run TI methods. See https://docs.docker.com/install for instructions. For windows 10 you can install [Docker CE](https://store.docker.com/editions/community/docker-ce-desktop-windows), older windows installations require the [Docker toolbox](https://docs.docker.com/toolbox/overview/).

You can test whether docker is correctly installed by running:

```{r, message = TRUE, echo = TRUE}
dynwrap::test_docker_installation(detailed = TRUE)
```

This command will give helpful tips if some parts of the installation are missing.